---
title: "HJA_Other_Analyses"
author: "Mallard"
date: "4/27/2023"
output: html_document
---

This file is a grab bag of analyses done with the HJA data that don't easily fit into a stand alone file. Some of this is scratch, others didn't go anywhere. 

Filter and restructure q data for WS1:
- Total Q (in inches), rather than mean, max, min
- Convert DATE from character to date format
- Convert inches to mm
```{r WS1q}
qWS1 <- qRaw %>% 
  filter(SITECODE == "GSWS01") %>% 
  select(DATE,TOTAL_Q_AREA) %>% 
  rename(runoff = TOTAL_Q_AREA) %>% 
  mutate(DATE = mdy(DATE)) %>% 
  mutate(runoff = runoff * 25.4)

# This clunky line is dealing with the fact that ymd functions set default 2000 cutoff year at 1968. Might be a way to integrate this operation with piped operation above
qWS1$DATE[year(qWS1$DATE)>2022] <- qWS1$DATE[year(qWS1$DATE)>2022] - years(100) 

```

Combine q and met data:
- Reduce to common days
- Filter to after WY2000
- Add water year column
```{r WS1data}
WS1data <- merge(met,qWS1,by = 'DATE')

WS1data <- WS1data[WS1data$DATE > as.Date('1999-09-30'),] %>% 
  mutate(yr = year(DATE), mn = month(DATE)) %>% 
  mutate(WY = ifelse(mn > 9, yr + 1, yr)) %>% 
  select(-c(yr,mn))
```

Make some summary datasets:
-Annual
-Monthly
```{r WS1summaries}
WS1annual <- WS1data %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(runoff),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip,
         ET = precip - runoff,
         er = ET/precip) %>% 
  ungroup()

WS1monthly <- WS1data %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(runoff),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

# WS1months <- WS1monthly %>% 
#   group_by(month) %>% 
#   summarize(runoff = mean(runoff),
#             precip = mean(precip),
#             rr = mean(rr),
#             rrsd = sd(rr)) %>% 
#   summarize(runoffsd = sd(runoff),
#             precipsd = sd(precip),
#             rrsd = sd(rr)) %>% 
#   ungroup
```

Make some plots:
- Annual runoff, precip, rr
- Monthly rr

```{r WS1figs}
#Annual
annPr <- ggplot(data = WS1annual, aes(x=WY,y=precip)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
annR <- ggplot(data = WS1annual, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'coral')
annRR <- ggplot(data = WS1annual, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkolivegreen3')

PrRR <- ggplot(data = WS1annual, aes(x=precip,y=rr)) +
  geom_point(size=5)
PrER <- ggplot(data = WS1annual, aes(x=precip,y=er)) +
  geom_point(size=5)

annPr
annR
annRR
PrRR
PrER

#Monthly
monRr <- ggplot(data = WS1monthly, aes(x=DATE,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkolivegreen3')
monRrBox <- ggplot(data = WS1monthly, aes(x=month, y = rr)) +
  geom_boxplot()

monRr
monRrBox
```

Make some figures with this larger dataset (allData):
- Runoff annual
- Runoff ratio boxplots
- Runoff ratio annual values
```{r allWSfigs}
#Annual RR and precip for each WS
#WS1
WS1 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS1),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS1Q <- ggplot(WS1, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS1RR <- ggplot(WS1, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS2
WS2 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS2),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS2Q <- ggplot(WS2, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS2RR <- ggplot(WS2, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS3
WS3 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS3, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS3Q <- ggplot(WS3, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS3RR <- ggplot(WS3, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS10
WS10 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS10),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS10Q <- ggplot(WS10, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS10RR <- ggplot(WS10, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#Boxplots of monthly RR for each 
#WS1
WS1mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS1, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS1monRR <- ggplot(data = WS1mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS2
WS2mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS2, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS2monRR <- ggplot(data = WS2mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS3
WS3mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS3, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS3monRR <- ggplot(data = WS3mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS10
WS10mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS10, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS10monRR <- ggplot(data = WS10mon, aes(x=month, y = rr)) +
  geom_boxplot()

#Print these plots
WS1Q
WS1RR
WS2Q
WS2RR
WS3Q
WS3RR
WS10Q
WS10RR

WS1monRR
WS2monRR
WS3monRR
WS10monRR
```

Use allData to redo dataset for ws2 and add 7day mean low flow. Create annual time series from it
```{r ws2LowFlow}
#Find running 7 day mean
WS2data <- allData %>% 
  select(DATE, precip, WS2, WY) %>% 
  rename(Q = WS2) %>% 
  mutate(Low7Day = rollmean(Q,
                            7,
                            fill = NA,
                            na.rm = FALSE,
                            align = "center"))

# #Plot and subset to check data
WS2data %>%
  filter(DATE > mdy("4-01-1986") & DATE < mdy("11-01-1986")) %>%
  ggplot(aes(DATE, Q, color = "daily"))+
    geom_line()+
    geom_line(aes(x = DATE, y = Low7Day, color = "7 day mean"))

# testSubset<-WS2data %>%
#   filter(DATE > mdy("01-01-1987") & DATE < mdy("12-31-1987"))

#Create annual low flow, along with total precip and date of low flow, add precip for preceding WY
#Discard 1979 and 2018 for incomplete data
WS2LowFlow <- WS2data %>% 
  mutate(Year = year(DATE)) %>% 
  group_by(Year) %>% 
  summarize(minQ = min(Low7Day, na.rm = TRUE),
            minQDate = DATE[which.min(Low7Day)],
            annP = sum(precip, na.rm = TRUE)) %>% 
  mutate(minQDoy = yday(minQDate))

WS2WYprecip<- WS2data %>% 
  group_by(WY) %>% 
  summarize(annPWY = sum(precip, na.rm = TRUE))

WS2LowFlow$annPWY <- WS2WYprecip$annPWY

WS2LowFlow <- WS2LowFlow %>% 
  filter(Year != 1979 & Year != 2018)

#Some figures summarizing this annual dataset
WS2LowFlow %>% 
  ggplot(aes(y = minQ)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min daily runoff (mm)")
WS2LowFlow %>% 
  ggplot(aes(x = Year, y = minQ)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min daily runoff (mm)")

WS2LowFlow %>% 
  ggplot(aes(y = minQDoy)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min day of year (mm)") 
WS2LowFlow %>% 
  ggplot(aes(x = Year, y = minQDoy)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min day of year (mm)")

WS2LowFlow %>% 
  ggplot(aes(y = annPWY)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "annual precip (mm)")
WS2LowFlow %>% 
  ggplot(aes(x = Year, y = annPWY)) +
  geom_bar(stat = 'identity') +
  labs(y = "annual precip (mm)")
  
#Figures comparing them
# ggplot(data = WS2LowFlow, aes(x = annP, y = minQ))+
#   geom_point()+
#    labs(x = "annual precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS2LowFlow, aes(x = annPWY, y = minQ))+
  geom_point()+
   labs(x = "prior WY precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS2LowFlow, aes(x = minQDoy, y = minQ))+
  geom_point()+
  labs(x = "7 day min day of year (mm)", y = "7 day min daily runoff (mm)")

```

Do the same for WS1
```{r ws1LowFlow}
#Find running 7 day mean
WS1data <- allData %>% 
  select(DATE, precip, WS1, WY) %>% 
  rename(Q = WS1) %>% 
  mutate(Low7Day = rollmean(Q,
                            7,
                            fill = NA,
                            na.rm = FALSE,
                            align = "center"))

# #Plot and subset to check data
# WS1data %>%
#   filter(DATE > mdy("4-01-1986") & DATE < mdy("11-01-1986")) %>%
#   ggplot(aes(DATE, Q, color = "daily"))+
#     geom_line()+
#     geom_line(aes(x = DATE, y = Low7Day, color = "7 day mean"))

# testSubset<-WS2data %>%
#   filter(DATE > mdy("01-01-1987") & DATE < mdy("12-31-1987"))

#Create annual low flow, along with total precip and date of low flow, add precip for preceding WY
#Discard 1979 and 2018 for incomplete data
WS1LowFlow <- WS1data %>% 
  mutate(Year = year(DATE)) %>% 
  group_by(Year) %>% 
  summarize(minQ = min(Low7Day, na.rm = TRUE),
            minQDate = DATE[which.min(Low7Day)],
            annP = sum(precip, na.rm = TRUE)) %>% 
  mutate(minQDoy = yday(minQDate))

WS1WYprecip<- WS1data %>% 
  group_by(WY) %>% 
  summarize(annPWY = sum(precip, na.rm = TRUE))

WS1LowFlow$annPWY <- WS1WYprecip$annPWY

WS1LowFlow <- WS1LowFlow %>% 
  filter(Year != 1979 & Year != 2018)

#Some figures summarizing this annual dataset
WS1LowFlow %>% 
  ggplot(aes(y = minQ)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min daily runoff (mm)")
WS1LowFlow %>% 
  ggplot(aes(x = Year, y = minQ)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min daily runoff (mm)")

WS1LowFlow %>% 
  ggplot(aes(y = minQDoy)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min day of year (mm)") 
WS1LowFlow %>% 
  ggplot(aes(x = Year, y = minQDoy)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min day of year (mm)")

WS1LowFlow %>% 
  ggplot(aes(y = annPWY)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "annual precip (mm)")
WS1LowFlow %>% 
  ggplot(aes(x = Year, y = annPWY)) +
  geom_bar(stat = 'identity') +
  labs(y = "annual precip (mm)")
  
#Figures comparing them
# ggplot(data = WS2LowFlow, aes(x = annP, y = minQ))+
#   geom_point()+
#    labs(x = "annual precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS1LowFlow, aes(x = annPWY, y = minQ))+
  geom_point()+
   labs(x = "prior WY precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS1LowFlow, aes(x = minQDoy, y = minQ))+
  geom_point()+
  labs(x = "7 day min day of year (mm)", y = "7 day min daily runoff (mm)")
```
Calculate a 7-day low flow for all days and the same day every year in WS1,2,3,10. Same day is 7-day window centered on 8/30, doy = 242. 
Should add code to dynamically generate variable names from WS names!
```{r WS2LowFlowAug}
WSLowFlowData <- allData %>% 
  select(DATE, precip, WS1, WS2, WS3, WS10, meanVpd, meanTemp, WY) %>% 
  mutate(WS1Low = rollmean(WS1, 7, fill = NA, na.rm = FALSE, align = "center"),
         WS2Low = rollmean(WS2, 7, fill = NA, na.rm = FALSE, align = "center"),
         WS3Low = rollmean(WS3, 7, fill = NA, na.rm = FALSE, align = "center"),
         WS10Low = rollmean(WS10, 7, fill = NA, na.rm = FALSE, align = "center"))

#Subsetting and plotting                            
# allData %>%
#   filter(DATE > mdy("1-01-1996") & DATE < mdy("12-01-1996")) %>%
#   ggplot(aes(DATE, WS3, color = "daily")) +
#   geom_line()

#Create annual aug low flow
#Discard before 1979 and after 2018 for incomplete data
WSLowFlowAnn <- WSLowFlowData %>% 
  mutate(Year = year(DATE)) %>% 
  group_by(Year) %>% 
  summarize(WS1minQAug = WS1Low[yday(DATE) == 242],
            WS2minQAug = WS2Low[yday(DATE) == 242],
            WS3minQAug = WS3Low[yday(DATE) == 242],
            WS10minQAug = WS10Low[yday(DATE) == 242]) %>% 
  filter(Year > 1979 & Year < 2019)

#Check where variables exist for WS2Data
imputeTS::ggplot_na_distribution(x = WS2data$meanVpd, x_axis_labels = WS2data$DATE)
imputeTS::ggplot_na_distribution(x = WS2data$meanTemp, x_axis_labels = WS2data$DATE)


#Plot all aug 30 low-flows
WSLowFlowAnn %>% 
  gather(WS,lowFlowAug,c("WS1minQAug","WS2minQAug","WS3minQAug","WS10minQAug")) %>% 
  ggplot(aes(x = Year, y = lowFlowAug, fill = WS)) + geom_bar(position = "dodge", stat = "identity") + theme_classic()

WSLowFlowAnn %>% 
  gather(WS,lowFlowAug,c("WS1minQAug","WS2minQAug","WS3minQAug","WS10minQAug")) %>% 
  ggplot(aes(x = Year, y = lowFlowAug, color = WS)) + geom_line() + theme_classic()

```


Exploratory analysis using WSXLowFlow data and climate variables with climwin package. Separate execution bc of runtime
- extract data for all WS low flow Aug 30
- slidingwin does not work with NAs in the predictive (climate) data. Replace those values with 0 for now
```{r ClimaWinData}
climData <- allData %>% 
  select(DATE,precip,meanTemp) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y"))

climData$precip[is.na(climData$precip)] <- 0
climData$meanTemp[is.na(climData$meanTemp)] <- 0

lowFlowData <- WSLowFlowAnn %>% 
  select(WS1minQAug, WS2minQAug, WS3minQAug, WS10minQAug, Year) %>% 
  mutate(txtDate = paste0("30/08/",Year)) #Add 30-08-XXXX formatted date
```

Execute climwin main function for multiple watersheds
```{r ClimaWinDataExec}
WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug","WS10minQAug")
climWinModels <- vector(mode = "list", length = length(WSvarnames)) #preallocate list

for (i in 1:length(WSvarnames)){
  lowFlowTemp <- lowFlowData %>% #Extract specific WS and remove rows with NA
    select(Year, txtDate, minQ = WSvarnames[i]) %>% 
    drop_na()
    
  winOut <- slidingwin(xvar = list(climData$meanTemp),
                     cdate = climData$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,10),
                     type = "relative",
                     stat = "mean",
                     func = "lin"
                     )

climWinModels[[i]] <- winOut[[1]]$Dataset
}
```

Visualize data tables and plots using built in functions. Added loop to plot multiples
```{r ClimaWinDataVis}
for (i in 1:length(WSvarnames)) {
  plot1 <- plotwin(dataset = climWinModels[[i]]) + ggtitle(paste0("Open/close for 95% set ", WSvarnames[i]))
  plot2 <- plotdelta(dataset = climWinModels[[i]]) + ggtitle(paste0("Model AICs for ", WSvarnames[i]))
  plot3 <- plotweights(dataset = climWinModels[[i]]) + ggtitle(paste0("95% set for ", WSvarnames[i]))
  print(plot1)
  print(plot2)
  print(plot3)
}

# head(climWinModels[[4]])
# 
# plotdelta(dataset = climWinModels[[4]])
# plotweights(dataset = climWinModels[[4]])

```

Calculating PET for HJA using PRIMET data
- Input data is daily temp
- Calculate estimated monthly PET from Thornwaite method: https://en.wikipedia.org/wiki/Potential_evaporation for example
  - PET = 16 * (L/12) * (N/30) * (10Td/I)^a
  - L: average day length of the month at that location
  - N: # of days in the month
  - Td: Average daily temperature in the month; 0 if temp is <0
  - a: (6.75e-7) * I^3 - (7.71e-5) * I^2 + (1.792e-2) * I + 0.49239
  - I (site heat index): sum((Tdi/5)^1.514) for i = 1:12
  - Tdi: Mean monthly temp for month i
  
```{r PET_Thorn}
#Setting up input values
hjaLat <- 44.23 #For daylength calcs

monthData <- allData %>% 
  select(DATE,meanTemp) %>% 
  mutate(doy = yday(DATE)) %>% #Create day of year variable
  mutate(dayLength = chillR::daylength(hjaLat,doy)[[3]]) %>% #Get daylength for each of these
  group_by(month = floor_date(DATE, "month")) %>% 
  summarise(meanMonTemp = mean(meanTemp, na.rm = "TRUE"),
            meanDaylength = mean(dayLength, na.rm = "TRUE")) %>% #Summarize into monthly means for temp and daylength
  na.omit %>% 
  mutate(meanMonTempTh = ifelse(meanMonTemp > 0, meanMonTemp, 0)) %>% #Create temp variable that is 0 if temp is below 0
  mutate(monthDays = as.numeric(days_in_month(month))) %>%  #Calculate days in month for each month
  mutate(monthYear = year(month))

#Find years that aren't 12 complete months bc heat index in Thornwaithe calculation requires 12 months
monthCounts <- monthData %>% 
  select(monthYear) %>% 
  mutate(monthsTemp = 1) %>% 
  group_by(monthYear) %>% 
  summarise(monthCounts = sum(monthsTemp))

incompleteYears <- monthCounts$monthYear[monthCounts$monthCounts != 12]

#Remove years from monthData that don't have 12 months
monthData <- monthData %>% 
  filter(!(monthYear %in% incompleteYears))

#Calculate heat index I for each month, sum to yearly heat index, calculate exponent alpha
thornHeatI <- monthData %>% 
  select(month, monthYear, meanMonTempTh) %>% 
  mutate(monI = (meanMonTempTh/5)^1.514) %>% 
  group_by(monthYear) %>% 
  summarise(heatI = sum(monI)) %>% 
  mutate(aThorn = (6.75e-7 * heatI^3) - (7.71e-5 * heatI^2) + (0.01792 * heatI) + 0.49239)

#Join thornHeatI to monthData so there's a yearly heat index value for each month and calculate monthly/daily PET
thornPET <- left_join(monthData, thornHeatI, by = "monthYear") %>% 
  mutate(thornMonPET = 16 * (meanDaylength / 12) * (monthDays / 30) * (10 * meanMonTempTh / heatI)^aThorn) %>% 
  mutate(thornDayPET = thornMonPET / monthDays) %>% 
  select(month,monthDays,thornMonPET,thornDayPET)
```

Then, use PET to calculate storage in WS2
- Input data is daily/monthly PET, daily runoff, and daily precip
- Make monthly summaries of Q and P
- Water balance eqn: deltaS = P - ET - Q
- Assume PET ~= ET bc energy limited at HJA

```{r WS2_Storage}
#Create WS2 daily dataset from allData and thornPET
WS2Storage <- allData %>% 
  select(DATE,precip,WS2,WY) %>% 
  rename(runoff = WS2) %>% 
  na.omit() %>% 
  mutate(dateMonth = floor_date(DATE,"month")) %>% 
  left_join(., thornPET, by = c("dateMonth" = "month")) %>% 
  mutate(delS = precip - runoff - thornDayPET) %>% 
  mutate(dayS = cumsum(delS))
```

Some figs of storage in WS2
```{r WS2_storage_figs}
ggplot(data = WS2Storage, aes(DATE,dayS)) +
  geom_line()

ggplot(data = WS2Storage, aes(DATE,delS)) +
  geom_point()

WS2StorageAnnual <- WS2Storage %>% 
  group_by(WY) %>% 
  summarise(yearlyPET = sum(thornDayPET, na.rm = T), 
            yearlyRunoff = sum(runoff, na.rm = T),
            yearlyPrecip = sum(precip, na.rm = T))

ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = yearlyPET), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = yearlyRunoff), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = yearlyPrecip), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = (yearlyPET+yearlyRunoff)), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = (yearlyPrecip-yearlyRunoff)), stat = "identity") + scale_y_continuous(limits = c(0,3000))
```

Troubleshooting of primet data to explain odd figures in previous chunk
```{r PRIMET_troubleshooting}
#First, let's check for duplicate data in allData data frame
# tsAllData <- allData %>% 
#   mutate(tsYear = year(DATE)) %>% 
#   count(tsYear)

#Starting in 1994 amount of data increased by almost 10 fold


#Check some of the variables for multiple options
unique(tsRawPrecipData$PRECIP_METHOD)
unique(tsRawPrecipData$PROBE_CODE)

#Temp
tsRawTempData <- tempRaw %>% 
  filter(SITECODE == "PRIMET") %>% 
  mutate(DATE = as_date(DATE))

#Check some of the variables for multiple options
unique(tsRawTempData$AIRTEMP_METHOD)
unique(tsRawTempData$PROBE_CODE)

#Multiple values for all of these. Check in raw data where duplicate days coming from
#Filter to 1994 in both data
tsRawPrecipDataSub <- tsRawPrecipData %>% 
  filter(between(DATE, as.Date('1994-01-01'), as.Date('1994-12-31')))

tsRawTempDataSub <- tsRawTempData %>% 
  filter(between(DATE, as.Date('1980-01-01'), as.Date('1980-12-31')))


#For temp, what heights are available through the record
tempHeights <- tsRawTempData %>% 
  select(HEIGHT,DATE) %>% 
  mutate(myYears = year(DATE)) %>% 
  filter(HEIGHT == 150) %>% 
  group_by(myYears) %>% 
  count(HEIGHT)


unique(tempHeights$HEIGHT)

#Ok, so filtering temp data to 150 m only. Now figure out what other unique values
# tsRawTempData <- tempRaw %>% 
#   filter(SITECODE == "PRIMET") %>% 
#   mutate(DATE = as_date(DATE)) %>% 
#   filter(HEIGHT == 150)
# 
# unique(tsRawTempData$AIRTEMP_METHOD)
# unique(tsRawTempData$PROBE_CODE)
# 
# sensorExtent <- ggplot(data = tsRawTempData) + 
#   geom_point(aes(x = DATE, y = as.factor(PROBE_CODE)))
# 
# sensorExtent
# 
# ggplotly(sensorExtent)

#Filtering additionally to sensors 5,6, and a few days of 4
tsRawTempData <- tempRaw %>% 
  filter(SITECODE == "PRIMET") %>% 
  mutate(DATE = as_date(DATE)) %>% 
  filter(HEIGHT == 150) %>% 
  filter(!(PROBE_CODE == "AIRPRI03")) %>% 
  filter(PROBE_CODE %in% c("AIRPRI06", "AIRPRI05") | 
           (PROBE_CODE == "AIRPRI04" & between(DATE, as.Date("2004-6-15"), as.Date("2004-6-25"))))

sensorExtent <- ggplot(data = tsRawTempData) + 
  geom_point(aes(x = DATE, y = as.factor(PROBE_CODE)))

sensorExtent

ggplotly(sensorExtent)

#Filtering to methods
unique(tsRawTempData$AIRTEMP_METHOD)
unique(tsRawTempData$PROBE_CODE)

#Method time extents
sensorExtent <- ggplot(data = tsRawTempData) +
  geom_point(aes(x = DATE, y = as.factor(AIRTEMP_METHOD)))

sensorExtent

ggplotly(sensorExtent)

#Similar work on filtering precip data
tsRawPrecipData <- precipRaw %>% 
  filter(SITECODE == "PRIMET") %>% 
  mutate(DATE = as_date(DATE))

unique(tsRawPrecipData$PROBE_CODE)
unique(tsRawPrecipData$PRECIP_METHOD)

sensorExtent <- ggplot(data = tsRawPrecipData) +
  geom_point(aes(x = DATE, y = as.factor(PRECIP_METHOD)))

ggplotly(sensorExtent)

nrow(tsRawPrecipData)
length(unique(tsRawPrecipData$DATE))

#Same filtering with vpd data
tsRawVpdData <- vpdRaw %>% 
  filter(SITECODE == "PRIMET") %>% 
  mutate(DATE = as_date(DATE)) %>% 
  filter(HEIGHT == 150) %>% 
  filter((PROBE_CODE == "VPDPRI05") |
           ((PROBE_CODE == "VPDPRI04") & (DATE > as_date("2000-05-29")))
           )

unique(tsRawVpdData$PROBE_CODE)
unique(tsRawVpdData$VPD_METHOD)
unique(tsRawVpdData$HEIGHT)

sensorExtent <- ggplot(data = tsRawVpdData) +
  geom_point(aes(x = DATE, y = as.factor(PROBE_CODE)))

ggplotly(sensorExtent)

#Still to-do:
#Sort out flags on data to remove low-quality measurements
#Look through discharge data to do the same


```

