---
title: "HJA_ET_fromMoore2004"
output: html_notebook
---

Attempts to leverage data from Moore et al 2004 study on transpiration at HJA WS1 and 2

Load data from digitized csvs
```{r Data}
Figure3aData <- read.csv("Data/Moore2004Data/Figure3a.csv")
Figure3bData <- read.csv("Data/Moore2004Data/Figure3b.csv")
Figure7Data <- read.csv("Data/Moore2004Data/Figure7.csv")
```

Restructure data:
- Make three df (sap flux 1999 and 2000 and transpiration)
- Lots of fiddling to get day of year into the right date
- Aggregate into daily values
- Linear interpolation of missing values

```{r structureData}
sapFlux1999 <- Figure3bData %>% 
  rename(sFlux = SapFluxDens) %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, tree) %>% 
  summarise(sFluxDaily = mean(sFlux)) %>% 
  ungroup() %>% 
  spread(tree,sFluxDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(PSMEy = na.approx(PSMEy),
         ALRUy = na.approx(ALRUy)) %>% 
  gather(tree, sFlux, PSMEy, ALRUy)
  

sapFlux2000 <- Figure3aData %>% 
  rename(sFlux = SapFluxDens) %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, tree) %>% 
  summarise(sFluxDaily = mean(sFlux)) %>% 
  ungroup() %>% 
  spread(tree,sFluxDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(PSMEy = na.approx(PSMEy),
         PSMEo = na.approx(PSMEo)) %>% 
  gather(tree, sFlux, PSMEy, PSMEo)
  
trans2000 <- Figure7Data %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, tree) %>% 
  summarise(transDaily = mean(trans)) %>% 
  ungroup() %>% 
  spread(tree,transDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(PSMEy = na.approx(PSMEy),
         PSMEo = na.approx(PSMEo)) %>% 
  gather(tree, trans, PSMEy, PSMEo)

#Plot these dfs to check how they match with Moore 2004
Figure3aMoore <- sapFlux2000 %>% 
  ggplot(aes(x = DATE, y = sFlux)) +
  geom_line(aes(color = tree))
print(Figure3aMoore)

Figure3bMoore <- sapFlux1999 %>%
ggplot(aes(x = DATE, y = sFlux)) +
  geom_line(aes(color = tree))
print(Figure3bMoore)

Figure7Moore <- trans2000 %>% 
  ggplot(aes(x = DATE, y = trans)) +
  geom_line(aes(color = tree))
print(Figure7Moore)

```

Now get vpd and radiation data for these same time periods but also as early as march or so to back extrapolate
```{r metData4Trans}
vpdMoore <- allData %>% 
  filter(between(DATE, as.Date("1999-03-01"), as.Date("2000-11-01"))) %>% 
  select(DATE,meanVpd) %>% 
  ggplot(aes(x = DATE, y = meanVpd)) + geom_line()

ggplotly(vpdMoore)

#Missing data from 2000-6-4 - 2000-6-21 and 2000-5-24 - 2000-5-31. Check if there are other data in this range at primet
vpdCheck <- vpdRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE,VPD_MEAN_DAY,SITECODE,HEIGHT,PROBE_CODE) %>% 
  filter(between(DATE, as.Date("2000-05-23"), as.Date("2000-06-01")) |
         between(DATE, as.Date("2000-06-03"), as.Date("2000-06-22"))) %>% 
  filter(!is.na(VPD_MEAN_DAY)) %>% 
  filter(SITECODE %in% c("PRIMET", "CS2MET"))

#Find radiation data for this period
radRaw <- read.csv("Data/MS00105_v6.csv")

metRad <- radRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE, SOLAR_TOT_DAY, SITECODE, HEIGHT, PROBE_CODE) %>% 
  filter(between(DATE, as.Date("1999-03-01"), as.Date("2000-11-01"))) %>% 
  filter(SITECODE == "PRIMET")

radPlot <- metRad %>% 
  ggplot(aes(x = DATE, y = SOLAR_TOT_DAY)) + geom_line()

ggplotly(radPlot)
#Data gap at 2000-5-24 - 2000-5-31
```

Check to see what other sensors and stations could be suitable for linear regression interpolation.
  Requires radRaw loaded in previous chunk from MS00105_v6.csv.
  First select a data range and create vpd and rad dfs
```{r gapFillingVpdRad1}
fromToDates <- c("1999-03-01", "2000-11-01")

#Vpd data from all sites, sensors, heights
vpdCheck <- vpdRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE,VPD_MEAN_DAY,SITECODE,HEIGHT,PROBE_CODE) %>% 
  filter(between(DATE, as.Date(fromToDates[1]), as.Date(fromToDates[2])))

#Radiation data from all sites, sensors, heights
radCheck <- radRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE,SOLAR_TOT_DAY,SITECODE,HEIGHT,PROBE_CODE) %>% 
  filter(between(DATE, as.Date(fromToDates[1]), as.Date(fromToDates[2])))
```

  Then get unique combos of site and height and test relationships between each of these and primet data (probe 04, not 05) 
```{r gapFillingVpdRad2}
varCheck <- vpdCheck
datesGap <- c("2000-05-25", "2000-05-30", "2000-06-05", "2000-06-20") #gaps for VPD
# datesGap <- c("2000-05-25", "2000-05-30") #gaps for radiation

uniqueCombos <- unique(varCheck[ , c("SITECODE","HEIGHT")]) %>% 
  filter(SITECODE != "PRIMET") %>% 
  mutate(numNAs = 0)

for (i in 1:nrow(uniqueCombos)) { #:nrow(uniqueCombos)
  varPRI <- varCheck %>% 
    filter(SITECODE == "PRIMET") %>% 
    filter(PROBE_CODE == "VPDPRI04" |
             PROBE_CODE == "RADPRI01")
  varOther <- varCheck %>% 
    filter(SITECODE == uniqueCombos$SITECODE[i]) %>% 
    filter(HEIGHT == uniqueCombos$HEIGHT[i])
  
  #Check how many NAs are in potential predictor data
  varOtherCheck <- varOther %>% 
    filter(between(DATE, as.Date(datesGap[1]), as.Date(datesGap[2])) |
           between(DATE, as.Date(datesGap[3]), as.Date(datesGap[4])))
  
  uniqueCombos$numNAs[i] <- sum(is.na(varOtherCheck$VPD_MEAN_DAY))
  if(uniqueCombos$numNAs[i] != 0) next
  
  #Plot bivariate of predictor and PRIMET data
  tempPlot <- left_join(varPRI, varOther, by = "DATE") %>% 
    select(DATE, 2, 6) %>% 
    rename(PRIvar = 2, OtherVar = 3) %>% 
    drop_na() %>% 
    ggplot(aes(x = OtherVar, y = PRIvar)) + geom_point() +
    geom_smooth(formula = y ~ x, method = "lm") +
    stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label") +
    labs(title = paste(uniqueCombos$SITECODE[i], uniqueCombos$HEIGHT[i], "cm"))
  print(tempPlot)
}
```

Finally use the best models to fill gaps in PRIMET data and save as a single data frame
  Best Models:
  - VPD: CS2MET, 150 cm
  - Radiation: CENMET, 625 cm
  Requires vpdCheck and radCheck from two chunks up
```{r gapFillingVpdRad3}
#Put PRIMET radiation, vpd in same df as predictor variables
PRIgfVpd <- vpdCheck %>% 
  filter(SITECODE %in% c("PRIMET", "CS2MET")) %>% 
  filter(PROBE_CODE != "VPDPRI05") %>% 
  select(DATE, VPD_MEAN_DAY, SITECODE) %>% 
  spread(SITECODE, VPD_MEAN_DAY) 

#Find lm, add prediction column, replace NAs in PRIMET with prediction
vpdReg <- lm(PRIMET ~ CS2MET, data = PRIgfVpd)

PRIgfVpd <- PRIgfVpd %>% 
  add_predictions(vpdReg) %>% 
  mutate(PRIMET = ifelse(is.na(PRIMET), pred, PRIMET))

#Plot data highlighting gap-fill
vpdGfPlot <- vpdCheck %>% 
  filter(SITECODE == "CS2MET") %>% 
  filter(HEIGHT == 150) %>% 
  ggplot(aes(x = DATE, y = VPD_MEAN_DAY)) + geom_line()
print(vpdGfPlot)

#Combine gap-filled vpd and radiation data into single df
```












