---
title: "HJA_Other_Analyses"
author: "Mallard"
date: "4/27/2023"
output: html_document
---

This file is a grab bag of analyses done with the HJA data that don't easily fit into a stand alone file. Some of this is scratch, others didn't go anywhere. 

Filter and restructure q data for WS1:
- Total Q (in inches), rather than mean, max, min
- Convert DATE from character to date format
- Convert inches to mm
```{r WS1q}
qWS1 <- qRaw %>% 
  filter(SITECODE == "GSWS01") %>% 
  select(DATE,TOTAL_Q_AREA) %>% 
  rename(runoff = TOTAL_Q_AREA) %>% 
  mutate(DATE = mdy(DATE)) %>% 
  mutate(runoff = runoff * 25.4)

# This clunky line is dealing with the fact that ymd functions set default 2000 cutoff year at 1968. Might be a way to integrate this operation with piped operation above
qWS1$DATE[year(qWS1$DATE)>2022] <- qWS1$DATE[year(qWS1$DATE)>2022] - years(100) 

```

Combine q and met data:
- Reduce to common days
- Filter to after WY2000
- Add water year column
```{r WS1data}
WS1data <- merge(met,qWS1,by = 'DATE')

WS1data <- WS1data[WS1data$DATE > as.Date('1999-09-30'),] %>% 
  mutate(yr = year(DATE), mn = month(DATE)) %>% 
  mutate(WY = ifelse(mn > 9, yr + 1, yr)) %>% 
  select(-c(yr,mn))
```

Make some summary datasets:
-Annual
-Monthly
```{r WS1summaries}
WS1annual <- WS1data %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(runoff),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip,
         ET = precip - runoff,
         er = ET/precip) %>% 
  ungroup()

WS1monthly <- WS1data %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(runoff),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

# WS1months <- WS1monthly %>% 
#   group_by(month) %>% 
#   summarize(runoff = mean(runoff),
#             precip = mean(precip),
#             rr = mean(rr),
#             rrsd = sd(rr)) %>% 
#   summarize(runoffsd = sd(runoff),
#             precipsd = sd(precip),
#             rrsd = sd(rr)) %>% 
#   ungroup
```

Make some plots:
- Annual runoff, precip, rr
- Monthly rr

```{r WS1figs}
#Annual
annPr <- ggplot(data = WS1annual, aes(x=WY,y=precip)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
annR <- ggplot(data = WS1annual, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'coral')
annRR <- ggplot(data = WS1annual, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkolivegreen3')

PrRR <- ggplot(data = WS1annual, aes(x=precip,y=rr)) +
  geom_point(size=5)
PrER <- ggplot(data = WS1annual, aes(x=precip,y=er)) +
  geom_point(size=5)

annPr
annR
annRR
PrRR
PrER

#Monthly
monRr <- ggplot(data = WS1monthly, aes(x=DATE,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkolivegreen3')
monRrBox <- ggplot(data = WS1monthly, aes(x=month, y = rr)) +
  geom_boxplot()

monRr
monRrBox
```

Make some figures with this larger dataset (allData):
- Runoff annual
- Runoff ratio boxplots
- Runoff ratio annual values
```{r allWSfigs}
#Annual RR and precip for each WS
#WS1
WS1 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS1),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS1Q <- ggplot(WS1, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS1RR <- ggplot(WS1, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS2
WS2 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS2),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS2Q <- ggplot(WS2, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS2RR <- ggplot(WS2, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS3
WS3 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS3, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS3Q <- ggplot(WS3, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS3RR <- ggplot(WS3, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS10
WS10 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS10),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS10Q <- ggplot(WS10, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS10RR <- ggplot(WS10, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#Boxplots of monthly RR for each 
#WS1
WS1mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS1, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS1monRR <- ggplot(data = WS1mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS2
WS2mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS2, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS2monRR <- ggplot(data = WS2mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS3
WS3mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS3, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS3monRR <- ggplot(data = WS3mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS10
WS10mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS10, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS10monRR <- ggplot(data = WS10mon, aes(x=month, y = rr)) +
  geom_boxplot()

#Print these plots
WS1Q
WS1RR
WS2Q
WS2RR
WS3Q
WS3RR
WS10Q
WS10RR

WS1monRR
WS2monRR
WS3monRR
WS10monRR
```


Calculating PET for HJA using PRIMET data
- Input data is daily temp
- Calculate estimated monthly PET from Thornwaite method: https://en.wikipedia.org/wiki/Potential_evaporation for example
  - PET = 16 * (L/12) * (N/30) * (10Td/I)^a
  - L: average day length of the month at that location
  - N: # of days in the month
  - Td: Average daily temperature in the month; 0 if temp is <0
  - a: (6.75e-7) * I^3 - (7.71e-5) * I^2 + (1.792e-2) * I + 0.49239
  - I (site heat index): sum((Tdi/5)^1.514) for i = 1:12
  - Tdi: Mean monthly temp for month i
  
```{r PET_Thorn}
#Setting up input values
hjaLat <- 44.23 #For daylength calcs

monthData <- allData %>% 
  select(DATE,meanTemp) %>% 
  mutate(doy = yday(DATE)) %>% #Create day of year variable
  mutate(dayLength = chillR::daylength(hjaLat,doy)[[3]]) %>% #Get daylength for each of these
  group_by(month = floor_date(DATE, "month")) %>% 
  summarise(meanMonTemp = mean(meanTemp, na.rm = "TRUE"),
            meanDaylength = mean(dayLength, na.rm = "TRUE")) %>% #Summarize into monthly means for temp and daylength
  na.omit %>% 
  mutate(meanMonTempTh = ifelse(meanMonTemp > 0, meanMonTemp, 0)) %>% #Create temp variable that is 0 if temp is below 0
  mutate(monthDays = as.numeric(days_in_month(month))) %>%  #Calculate days in month for each month
  mutate(monthYear = year(month))

#Find years that aren't 12 complete months bc heat index in Thornwaithe calculation requires 12 months
monthCounts <- monthData %>% 
  select(monthYear) %>% 
  mutate(monthsTemp = 1) %>% 
  group_by(monthYear) %>% 
  summarise(monthCounts = sum(monthsTemp))

incompleteYears <- monthCounts$monthYear[monthCounts$monthCounts != 12]

#Remove years from monthData that don't have 12 months
monthData <- monthData %>% 
  filter(!(monthYear %in% incompleteYears))

#Calculate heat index I for each month, sum to yearly heat index, calculate exponent alpha
thornHeatI <- monthData %>% 
  select(month, monthYear, meanMonTempTh) %>% 
  mutate(monI = (meanMonTempTh/5)^1.514) %>% 
  group_by(monthYear) %>% 
  summarise(heatI = sum(monI)) %>% 
  mutate(aThorn = (6.75e-7 * heatI^3) - (7.71e-5 * heatI^2) + (0.01792 * heatI) + 0.49239)

#Join thornHeatI to monthData so there's a yearly heat index value for each month and calculate monthly/daily PET
thornPET <- left_join(monthData, thornHeatI, by = "monthYear") %>% 
  mutate(thornMonPET = 16 * (meanDaylength / 12) * (monthDays / 30) * (10 * meanMonTempTh / heatI)^aThorn) %>% 
  mutate(thornDayPET = thornMonPET / monthDays) %>% 
  select(month,monthDays,thornMonPET,thornDayPET)
```

Then, use PET to calculate storage in WS2
- Input data is daily/monthly PET, daily runoff, and daily precip
- Make monthly summaries of Q and P
- Water balance eqn: deltaS = P - ET - Q
- Assume PET ~= ET bc energy limited at HJA

```{r WS2_Storage}
#Create WS2 daily dataset from allData and thornPET
WS2Storage <- allData %>% 
  select(DATE,precip,WS2,WY) %>% 
  rename(runoff = WS2) %>% 
  na.omit() %>% 
  mutate(dateMonth = floor_date(DATE,"month")) %>% 
  left_join(., thornPET, by = c("dateMonth" = "month")) %>% 
  mutate(delS = precip - runoff - thornDayPET) %>% 
  mutate(dayS = cumsum(delS))
```

Some figs of storage in WS2
```{r WS2_storage_figs}
ggplot(data = WS2Storage, aes(DATE,dayS)) +
  geom_line()

ggplot(data = WS2Storage, aes(DATE,delS)) +
  geom_point()

WS2StorageAnnual <- WS2Storage %>% 
  group_by(WY) %>% 
  summarise(yearlyPET = sum(thornDayPET, na.rm = T), 
            yearlyRunoff = sum(runoff, na.rm = T),
            yearlyPrecip = sum(precip, na.rm = T))

ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = yearlyPET), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = yearlyRunoff), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = yearlyPrecip), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = (yearlyPET+yearlyRunoff)), stat = "identity") + scale_y_continuous(limits = c(0,3000))
ggplot(data = WS2StorageAnnual) + geom_bar(aes(x = WY, y = (yearlyPrecip-yearlyRunoff)), stat = "identity") + scale_y_continuous(limits = c(0,3000))
```
