---
title: "HJAClimWin"
output: html_notebook
---

Requires low flow data and met data from "all data"

Exploratory analysis using WSXLowFlow data and climate variables with climwin package. Separate execution bc of runtime
- extract data for all WS low flow Aug 30
- slidingwin does not work with NAs in the predictive (climate) data. Replace those values with 0 for now in precip
```{r ClimWinData}
climData <- allData %>% 
  select(DATE, precip, meanTemp, meanVpd) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y")) %>% #Format data for use with slidingwin function
  # filter(DATE > as.Date("1978-04-01")) #For when PRIMET came online
  filter(DATE > as.Date("1988-07-12")) #For when vpd sensors came online

# climData$precip[is.na(climData$precip)] <- 0

lowFlowData <- WSLowFlowAnn %>% 
  select(WS1minQAug, WS2minQAug, WS3minQAug, WS10minQAug, Year) %>% 
  mutate(txtDate = paste0("30/08/",Year)) %>% 
  filter(Year > 1989) #For use with vpd
```

Checking on data provided by Karla (EDI_met_daily_daytime_nighttime...)
Make a similar climData df as above to plug into analyses
```{r KJdata}
load("Data/EDI_met_daily_daytime_nighttime_stats_20220505.Rdat")

ggplot_na_distribution(metnighttime_final$vpdmean, x_axis_labels = metdaytime_final$date)

climData <- metnighttime_final %>% 
  select(DATE = date, meanVpd = vpdmean) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y"))

```

Execute climwin main function for multiple watersheds
```{r ClimaWinDataExec}
# WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug","WS10minQAug")
WSvarnames <- "WS2minQAug"

climWinModels <- vector(mode = "list", length = length(WSvarnames)) #preallocate list

for (i in 1:length(WSvarnames)){
  lowFlowTemp <- lowFlowData %>% #Extract specific WS and remove rows with NA
    select(Year, txtDate, minQ = WSvarnames[i]) %>% 
    drop_na()
    
  winOut <- slidingwin(xvar = list(climData$meanVpd),
                     cdate = climData$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,0),
                     type = "relative",
                     stat = "mean",
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

climWinModels[[i]] <- winOut[[1]]$Dataset
}
```

Visualize data tables and plots using built in functions. Added loop to plot multiples
```{r ClimaWinDataVis}
for (i in 1:length(WSvarnames)) {
  plot1 <- plotwin(dataset = climWinModels[[i]]) + ggtitle(paste0("Open/close for 95% set ", WSvarnames[i]))
  plot2 <- plotdelta(dataset = climWinModels[[i]]) + ggtitle(paste0("Model AICs for ", WSvarnames[i]))
  plot3 <- plotweights(dataset = climWinModels[[i]])
  print(plot1)
  print(plot2)
  print(plot3)
}

# head(climWinModels[[4]])
# 
# plotdelta(dataset = climWinModels[[4]])
# plotweights(dataset = climWinModels[[4]])

```





