---
title: "HJAClimWin"
output: html_notebook
---

Requires low flow data and met data from "all data"

Exploratory analysis using WSXLowFlow data and climate variables with climwin package. Separate execution bc of runtime
- extract data for all WS low flow Aug 30
- slidingwin does not work with NAs in the predictive (climate) data. Replace those values with 0 for now in precip
```{r ClimWinData}
climDataPrecip <- allData %>% 
  select(DATE, precip, meanTemp, meanVpd) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y")) %>% #Format data for use with slidingwin function
  # filter(DATE > as.Date("1978-04-01")) #For when PRIMET came online
  filter(DATE > as.Date("1988-07-12")) #For when vpd sensors came online

climDataPrecip$precip[is.na(climDataPrecip$precip)] <- 0

lowFlowData <- WSLowFlowAnn %>% 
  ungroup() %>% 
  select(WS1minQAug, WS2minQAug, WS3minQAug, WS9minQAug, WS10minQAug, Year) %>% 
  mutate(txtDate = paste0("30/08/",Year)) %>% 
  filter(Year > 1989) %>%  #For use with vpd
  mutate(WS3minQAug = na_mean(WS3minQAug)) #Fill missing WS3 value in 1996 with mean
```

Checking on data provided by Karla (EDI_met_daily_daytime_nighttime...)
Make a climData df as above to plug into analyses for vpd
```{r KJdata}
load("Data/EDI_met_daily_daytime_nighttime_stats_20220505.Rdat")

# ggplot_na_distribution(metdaytime_final$vpdmean, x_axis_labels = metdaytime_final$date)

climDataVpd <- metdaytime_final %>% 
  select(DATE = date, meanVpd = vpdmean) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y"))

```

Execute climwin main function for multiple or single watersheds.
```{r ClimaWinDataExec}
# WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug", "WS9minQAug", "WS10minQAug")
WSvarnames <- "WS2minQAug"

climWinModels <- vector(mode = "list", length = length(WSvarnames)) #preallocate list

for (i in 1:length(WSvarnames)){
  lowFlowTemp <- lowFlowData %>% #Extract specific WS and remove rows with NA
    select(Year, txtDate, minQ = WSvarnames[i]) %>% 
    drop_na()
    
  winOut <- slidingwin(xvar = list(climData$meanVpd),
                     cdate = climData$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,10),
                     type = "relative",
                     stat = "mean",
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

climWinModels[[i]] <- winOut[[1]]$Dataset
}
```

Modified version of ClimaWinDataExec above to include runs for both precip and vpd and save in a single structure. Code in here to run randomization analysis, too
```{r ClimaWinDataExec2}
WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug", "WS9minQAug", "WS10minQAug")
# WSvarnames <- "WS2minQAug"

#preallocate lists
climWinModelsPrecip <- vector(mode = "list", length = length(WSvarnames)) 
climWinModelsVpd <- vector(mode = "list", length = length(WSvarnames)) 

# randWinModelsPrecip <- vector(mode = "list", length = length(WSvarnames)) 
# randWinModelsVpd <- vector(mode = "list", length = length(WSvarnames)) 

mm
for (i in 1:5){
  lowFlowTemp <- lowFlowData %>% #Extract specific WS and remove rows with NA
    select(Year, txtDate, minQ = WSvarnames[i]) %>% 
    drop_na()
  
  #Precip  
  winOut <- slidingwin(xvar = list(climDataPrecip$precip),
                     cdate = climDataPrecip$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,10),
                     type = "relative",
                     stat = "sum",
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

  climWinModelsPrecip[[i]] <- winOut[[1]]$Dataset
  
  # randOut <- randwin(repeats = 10, 
  #                    xvar = list(climDataPrecip$precip),
  #                    cdate = climDataPrecip$txtDate,
  #                    bdate = lowFlowTemp$txtDate,
  #                    baseline = lm(minQ ~ 1, data = lowFlowTemp),
  #                    cinterval = "day",
  #                    range = c(365,10),
  #                    type = "relative",
  #                    stat = "sum",
  #                    func = "lin",
  #                    cmissing = "method2" #This method replaces missing values with mean value of all records with same date
  #                    )
  # 
  # randWinModelsPrecip[[i]] <- randOut[[1]]
  
  #Vpd
  winOut <- slidingwin(xvar = list(climDataVpd$meanVpd),
                     cdate = climDataVpd$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,10),
                     type = "relative",
                     stat = "mean",
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

  climWinModelsVpd[[i]] <- winOut[[1]]$Dataset
  
  # randOut <- randwin(repeats = 10,
  #                    xvar = list(climDataVpd$meanVpd),
  #                    cdate = climDataVpd$txtDate,
  #                    bdate = lowFlowTemp$txtDate,
  #                    baseline = lm(minQ ~ 1, data = lowFlowTemp),
  #                    cinterval = "day",
  #                    range = c(365,10),
  #                    type = "relative",
  #                    stat = "mean",
  #                    func = "lin",
  #                    cmissing = "method2" #This method replaces missing values with mean value of all records with same date
  #                    )
  # 
  # randWinModelsVpd[[i]] <- randOut[[1]]
}

#Run randomization analyses for these ws's
# wsNum <- 1
# 
# pvalue(climWinModelsPrecip[[wsNum]], randWinModelsPrecip[[wsNum]], metric = "C", sample.size = 29)
# pvalue(climWinModelsVpd[[wsNum]], randWinModelsVpd[[wsNum]], metric = "C", sample.size = 29)
# 
# plothist(climWinModelsPrecip[[wsNum]], randWinModelsPrecip[[wsNum]])
# plothist(climWinModelsVpd[[wsNum]], randWinModelsVpd[[wsNum]])

save(climWinModelsPrecip, climWinModelsVpd, file = "climWinOutputs.RDat")
```

Visualize data tables and plots using built in functions. Added loop to plot multiples
```{r ClimaWinDataVis}
for (i in 1:length(WSvarnames)) {
  # plot1precip <- plotwin(dataset = climWinModels[[i]]) + ggtitle(paste0("Open/close for 95% set ", WSvarnames[i]))
  plot2precip <- plotdelta(dataset = climWinModelsPrecip[[i]]) + ggtitle(paste0("Precip model AICs for ", WSvarnames[i]))
  plot3precip <- plotweights(dataset = climWinModelsPrecip[[i]])
  # print(plot1precip)
  print(plot2precip)
  print(plot3precip)
  
  # plot1vpd <- plotwin(dataset = climWinModels[[i]]) + ggtitle(paste0("Open/close for 95% set ", WSvarnames[i]))
  plot2vpd <- plotdelta(dataset = climWinModelsVpd[[i]]) + ggtitle(paste0("Vpd model AICs for ", WSvarnames[i]))
  plot3vpd <- plotweights(dataset = climWinModelsVpd[[i]])
  # print(plot1vpd)
  print(plot2vpd)
  print(plot3vpd)
}


# head(climWinModels[[4]])
# 
# plotdelta(dataset = climWinModels[[4]])
# plotweights(dataset = climWinModels[[4]])

```

Look at individual model fits. Choose which model to plot based on best models output from slidingwin output and then use singlewin to get the specific model
```{r plotIndvModels}
winOpen <- 365
winClose <- 10
ws <- 1 #Which watershed: WS1, WS2, WS3, WS9, WS10
climVar <- "VPD" #Precip or VPD

WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug", "WS9minQAug", "WS10minQAug")
WSnames <- str_sub(WSvarnames, 1, -8)

#Select correct climate data
if (climVar == "Precip"){
  climDataTemp <- climDataPrecip %>% 
    select(txtDate, climPredictor = precip)
} else if (climVar == "VPD"){
  climDataTemp <- climDataVpd %>% 
    select(txtDate, climPredictor = meanVpd)
} else {
  stop("climVar must be precip vpd")
}

#Select correct watershed data
lowFlowDataTemp <- lowFlowData %>% 
  select(txtDate, minQAug = WSvarnames[ws])

#Select correct method for if it's precip (sum) or vpd (mean)
if (climVar == "Precip") {
  statType <- "sum"
} else if (climVar == "VPD") {
  statType <- "mean" 
} else {
  stop("climVar must be precip vpd")
}

#Get model to plot
model2plot <- singlewin(xvar = list(climDataTemp$climPredictor),
                     cdate = climDataTemp$txtDate,
                     bdate = lowFlowDataTemp$txtDate,
                     baseline = lm(minQAug ~ 1, data = lowFlowDataTemp),
                     cinterval = "day",
                     range = c(winOpen, winClose),
                     type = "relative",
                     stat = statType,
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

#Select correct output dataset for climate variable
if (climVar == "Precip") {
  climDataset <- climWinModelsPrecip[[ws]]
} else if (climVar == "VPD") {
  climDataset <- climWinModelsVpd[[ws]]
} else {
  stop("climVar must be 1 for precip or 2 for vpd")
}
  
plotbest(dataset = climDataset, 
         bestmodel = model2plot$BestModel,
         bestmodeldata = model2plot$BestModelData)

#Manual version of above plot
modelPlot <- ggplot(data = model2plot$BestModelData, aes(x = climate, y = yvar)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label") +
  labs(title = paste("Window", climVar, "vs Aug 30 LF for", WSnames[ws]),
       y = "7 day mean Q (mm/day)",
       x = "Total precip (mm)")
print(modelPlot)
```

Filter models to only those greater than some length of time and look at box plots and tile plots of top X open and close windows. Visualize these results with boxplots and tile plots.
```{r TopModels}
#inputs
winLength <- 7 #number of days of minimum window
numTopModels <- 20 #top number of models as ranked by AIC to plot

#Pre-allocate data frame
topClimModels <- data.frame(deltaAICc = double(),
                            WindowOpen = integer(),
                            WindowClose = integer(),
                            WSname = character(),
                            climPred = character())

for (i in 1:3) {
  #Precip models
  tempClimModels <- climWinModelsPrecip[[i]] %>% 
    select(deltaAICc, WindowOpen, WindowClose) %>% 
    mutate(WSname = str_sub(WSvarnames[i], 1, -8),
           climPred = "precip") %>% 
    mutate(lengthWin = WindowOpen - WindowClose + 1) %>% 
    filter(lengthWin >= winLength) %>% 
    select(-lengthWin) %>% 
    slice_min(deltaAICc, n = numTopModels)
  
  topClimModels <- rbind(topClimModels,tempClimModels)
  
  #Vpd models
  tempClimModels <- climWinModelsVpd[[i]] %>% 
    select(deltaAICc, WindowOpen, WindowClose) %>% 
    mutate(WSname = str_sub(WSvarnames[i], 1, -8),
           climPred = "vpd") %>% 
    mutate(lengthWin = WindowOpen - WindowClose + 1) %>% 
    filter(lengthWin >= winLength) %>% 
    select(-lengthWin) %>% 
    slice_min(deltaAICc, n = numTopModels)
  
  topClimModels <- rbind(topClimModels,tempClimModels)
} 

# :length(WSvarnames)

#Plot some boxplots of top models
boxTopModels <- topClimModels %>% 
  select(-deltaAICc) %>% 
  gather(winXO, day, WindowOpen:WindowClose) %>% 
  ggplot(aes(x = winDays2Dates(day), y = winXO)) +
  geom_boxplot() +
  facet_grid(WSname ~ climPred)

print(boxTopModels)

#Plot tile plots of top models
winDaysVec <- min(topClimModels$WindowClose):max(topClimModels$WindowOpen) #Create vector of all days in any window 

WSPredCombos <- unique(topClimModels[c("WSname", "climPred")])
WSCombos <- WSPredCombos$WSname
PredCombos <- WSPredCombos$climPred

tileTopModels <- data.frame(winDay = rep(winDaysVec, each = nrow(WSPredCombos))) %>%
  mutate(WS = rep(WSCombos, times = length(winDaysVec))) %>% 
  mutate(climVar = rep(PredCombos, times = length(winDaysVec))) %>% 
  mutate(nModDay = 0)

for (i in 1:nrow(tileTopModels)) {
  tempTopClimModels <- topClimModels %>% 
    filter(WSname == tileTopModels$WS[i] & 
             climPred == tileTopModels$climVar[i]) %>% 
    filter(WindowClose <= tileTopModels$winDay[i] &
             WindowOpen >= tileTopModels$winDay[i])
  
  tileTopModels$nModDay[i] <- nrow(tempTopClimModels)
}

tilePlots <- tileTopModels %>% 
  ggplot(aes(x = winDays2Dates(winDay), y = WS, fill = nModDay)) +
  geom_tile() + facet_wrap(~ climVar) +
  labs(title = paste("Top ", numTopModels, " models explaining Aug 30 low flow"),
       x = "Open and close days",
       y = "Watershed",
       fill = "# of models"
       )
print(tilePlots)

#Just WS2 as a bar chart
WS2WinPlot <- tileTopModels %>% 
  filter(WS == "WS2") %>% 
  mutate(winDate = winDays2Dates(winDay)) %>% 
  ggplot(aes(x = winDate, y = nModDay, fill = nModDay)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ climVar) +
  scale_y_continuous(expand = c(0,0))

print(WS2WinPlot)

```


This function is for converting window open and window close days to dates for plotting purposes
```{r days2datesBeforeAug30}
winDays2Dates <- function(d) {
  doys <- 241 - d #Convert day before Aug 30 to day of year. R starts from day 0 not 1
  output <- as.Date(doys, origin = "2023-01-01")
}
```














