---
title: "HJAdata"
author: "Mallard"
date: "10/24/2022"
output: html_document
---

Loading packages and input data
```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(climwin)
```

Input data catalog:
- HF00402: Flow data, daily runoff in
- MS00103: Precip data, mm
- HT00401: Temp @ gauges, deg C [per SteveW likely to be lower quality temp enclosures]
- MS00101: Daily (min/max/mean) Temp @ met stations, deg C
- MS00108: vpd (min/max/mean) @ met stations, mb
```{r inputData, echo=FALSE, message=FALSE}
qRaw <- read.csv("Data/HF00402_v12.csv")
metRaw <- read.csv("Data/MS00103_v8.csv")
tempRaw <- read.csv("Data/MS00101_v9.csv")
vpdRaw <- read.csv("Data/MS00108_v9.csv")

```
Daily precip values are in mm, daily q in inches (normalized to area)

Figuring out extents of data at met stations. 
```{r metDataExtent, echo=FALSE, message=FALSE}
metSiteLength <- metRaw %>% 
  group_by(SITECODE) %>% 
  summarise(
    timeStart = min(DATE),
    timeEnd = max(DATE)
  ) %>% 
  arrange(timeStart)

metSiteLength
```

Per SteveW PRIMET is the station to use, CS2MET is likely a bit too overgrown and may be undercatching.

Filter and restructure met:
- Only one site
```{r metData, echo=FALSE, message=FALSE}
met <- metRaw %>% 
  filter(SITECODE == "PRIMET") %>% 
  select(DATE,PRECIP_TOT_DAY) %>% 
  rename(precip = PRECIP_TOT_DAY) %>% 
  mutate(DATE = as_date(DATE))
```

Filter and restructure q data for ONE SITE:
- Total Q (in inches), rather than mean, max, min
- Convert DATE from character to date format
- Convert inches to mm
```{r WS1q}
qWS1 <- qRaw %>% 
  filter(SITECODE == "GSWS01") %>% 
  select(DATE,TOTAL_Q_AREA) %>% 
  rename(runoff = TOTAL_Q_AREA) %>% 
  mutate(DATE = mdy(DATE)) %>% 
  mutate(runoff = runoff * 25.4)

# This clunky line is dealing with the fact that ymd functions set default 2000 cutoff year at 1968. Might be a way to integrate this operation with piped operation above
qWS1$DATE[year(qWS1$DATE)>2022] <- qWS1$DATE[year(qWS1$DATE)>2022] - years(100) 

```

Combine q and met data:
- Reduce to common days
- Filter to after WY2000
- Add water year column
```{r WS1data}
WS1data <- merge(met,qWS1,by = 'DATE')

WS1data <- WS1data[WS1data$DATE > as.Date('1999-09-30'),] %>% 
  mutate(yr = year(DATE), mn = month(DATE)) %>% 
  mutate(WY = ifelse(mn > 9, yr + 1, yr)) %>% 
  select(-c(yr,mn))
```

Make some summary datasets:
-Annual
-Monthly
```{r WS1summaries}
WS1annual <- WS1data %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(runoff),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip,
         ET = precip - runoff,
         er = ET/precip) %>% 
  ungroup()

WS1monthly <- WS1data %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(runoff),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

# WS1months <- WS1monthly %>% 
#   group_by(month) %>% 
#   summarize(runoff = mean(runoff),
#             precip = mean(precip),
#             rr = mean(rr),
#             rrsd = sd(rr)) %>% 
#   summarize(runoffsd = sd(runoff),
#             precipsd = sd(precip),
#             rrsd = sd(rr)) %>% 
#   ungroup
```

Make some plots:
- Annual runoff, precip, rr
- Monthly rr

```{r WS1figs}
#Annual
annPr <- ggplot(data = WS1annual, aes(x=WY,y=precip)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
annR <- ggplot(data = WS1annual, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'coral')
annRR <- ggplot(data = WS1annual, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkolivegreen3')

PrRR <- ggplot(data = WS1annual, aes(x=precip,y=rr)) +
  geom_point(size=5)
PrER <- ggplot(data = WS1annual, aes(x=precip,y=er)) +
  geom_point(size=5)

annPr
annR
annRR
PrRR
PrER

#Monthly
monRr <- ggplot(data = WS1monthly, aes(x=DATE,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkolivegreen3')
monRrBox <- ggplot(data = WS1monthly, aes(x=month, y = rr)) +
  geom_boxplot()

monRr
monRrBox
```

Now lets put some datasets together for multiple watersheds (including WS1)
- WSs 1,2,3,10
- Total Q (in inches), rather than mean, max, min
- Convert DATE from character to date format
- Convert inches to mm
- Reformat as a wide table
- Set dates past 2022 back to 20th century (caused by 2 digit dates in raw data)
- Reduce to common days b/n met and q, but not all common days for watersheds
- Add water year column
- Merge with "met" variable from "metData" chunk to record length of met data
```{r WSdata}
#Q data
allQData <- qRaw %>% 
  filter(SITECODE %in% c("GSWS01","GSWS02","GSWS03","GSWS10")) %>% 
  select(DATE,TOTAL_Q_AREA,SITECODE) %>% 
  rename(runoff = TOTAL_Q_AREA) %>% 
  mutate(DATE = mdy(DATE)) %>% 
  mutate(runoff = runoff * 25.4) %>% 
  pivot_wider(names_from = SITECODE,values_from = runoff) %>% 
  rename(WS1 = GSWS01,
         WS2 = GSWS02,
         WS3 = GSWS03,
         WS10 = GSWS10) %>% 
  mutate(DATE = if_else(year(DATE)>2022,DATE-years(100),DATE)) #ymd function has trouble with 2 digit years
  
#Merge with met data
allData <- merge(met, allQData, by='DATE') %>%
  mutate(yr = year(DATE), mn = month(DATE)) %>%
  mutate(WY = ifelse(mn > 9, yr + 1, yr)) %>%
  select(-c(yr,mn))
```

Make some figures with this larger dataset:
- Runoff annual
- Runoff ratio boxplots
- Runoff ratio annual values
```{r allWSfigs}
#Annual RR and precip for each WS
#WS1
WS1 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS1),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS1Q <- ggplot(WS1, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS1RR <- ggplot(WS1, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS2
WS2 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS2),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS2Q <- ggplot(WS2, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS2RR <- ggplot(WS2, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS3
WS3 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS3, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS3Q <- ggplot(WS3, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS3RR <- ggplot(WS3, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#WS10
WS10 <- allData %>% 
  group_by(WY) %>% 
  summarize(runoff = sum(WS10),
            precip = sum(precip, na.rm = TRUE)) %>% 
  mutate(rr = runoff/precip)

WS10Q <- ggplot(WS10, aes(x=WY,y=runoff)) + 
  geom_bar(stat = 'identity', fill = 'steelblue')
WS10RR <- ggplot(WS10, aes(x=WY,y=rr)) + 
  geom_bar(stat = 'identity', fill = 'darkmagenta')

#Boxplots of monthly RR for each 
#WS1
WS1mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS1, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS1monRR <- ggplot(data = WS1mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS2
WS2mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS2, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS2monRR <- ggplot(data = WS2mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS3
WS3mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS3, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS3monRR <- ggplot(data = WS3mon, aes(x=month, y = rr)) +
  geom_boxplot()

#WS10
WS10mon <- allData %>% 
  group_by(DATE = floor_date(DATE, "month")) %>% 
  summarize(runoff = sum(WS10, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE)) %>%
  mutate(rr = runoff/precip) %>% 
  mutate(rr = ifelse(rr > 1, 1, rr)) %>% 
  mutate(month = factor(strftime(DATE,"%b"),levels=month.abb)) %>% 
  ungroup()

WS10monRR <- ggplot(data = WS10mon, aes(x=month, y = rr)) +
  geom_boxplot()

#Print these plots
WS1Q
WS1RR
WS2Q
WS2RR
WS3Q
WS3RR
WS10Q
WS10RR

WS1monRR
WS2monRR
WS3monRR
WS10monRR
```

Use allData to redo dataset for ws2 and add 7day mean low flow. Create annual time series from it
```{r ws2LowFlow}
#Find running 7 day mean
WS2data <- allData %>% 
  select(DATE, precip, WS2, WY) %>% 
  rename(Q = WS2) %>% 
  mutate(Low7Day = rollmean(Q,
                            7,
                            fill = NA,
                            na.rm = FALSE,
                            align = "center"))

# #Plot and subset to check data
WS2data %>%
  filter(DATE > mdy("4-01-1986") & DATE < mdy("11-01-1986")) %>%
  ggplot(aes(DATE, Q, color = "daily"))+
    geom_line()+
    geom_line(aes(x = DATE, y = Low7Day, color = "7 day mean"))

# testSubset<-WS2data %>%
#   filter(DATE > mdy("01-01-1987") & DATE < mdy("12-31-1987"))

#Create annual low flow, along with total precip and date of low flow, add precip for preceding WY
#Discard 1979 and 2018 for incomplete data
WS2LowFlow <- WS2data %>% 
  mutate(Year = year(DATE)) %>% 
  group_by(Year) %>% 
  summarize(minQ = min(Low7Day, na.rm = TRUE),
            minQDate = DATE[which.min(Low7Day)],
            annP = sum(precip, na.rm = TRUE)) %>% 
  mutate(minQDoy = yday(minQDate))

WS2WYprecip<- WS2data %>% 
  group_by(WY) %>% 
  summarize(annPWY = sum(precip, na.rm = TRUE))

WS2LowFlow$annPWY <- WS2WYprecip$annPWY

WS2LowFlow <- WS2LowFlow %>% 
  filter(Year != 1979 & Year != 2018)

#Some figures summarizing this annual dataset
WS2LowFlow %>% 
  ggplot(aes(y = minQ)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min daily runoff (mm)")
WS2LowFlow %>% 
  ggplot(aes(x = Year, y = minQ)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min daily runoff (mm)")

WS2LowFlow %>% 
  ggplot(aes(y = minQDoy)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min day of year (mm)") 
WS2LowFlow %>% 
  ggplot(aes(x = Year, y = minQDoy)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min day of year (mm)")

WS2LowFlow %>% 
  ggplot(aes(y = annPWY)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "annual precip (mm)")
WS2LowFlow %>% 
  ggplot(aes(x = Year, y = annPWY)) +
  geom_bar(stat = 'identity') +
  labs(y = "annual precip (mm)")
  
#Figures comparing them
# ggplot(data = WS2LowFlow, aes(x = annP, y = minQ))+
#   geom_point()+
#    labs(x = "annual precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS2LowFlow, aes(x = annPWY, y = minQ))+
  geom_point()+
   labs(x = "prior WY precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS2LowFlow, aes(x = minQDoy, y = minQ))+
  geom_point()+
  labs(x = "7 day min day of year (mm)", y = "7 day min daily runoff (mm)")

```

Do the same for WS1
```{r ws1LowFlow}
#Find running 7 day mean
WS1data <- allData %>% 
  select(DATE, precip, WS1, WY) %>% 
  rename(Q = WS1) %>% 
  mutate(Low7Day = rollmean(Q,
                            7,
                            fill = NA,
                            na.rm = FALSE,
                            align = "center"))

# #Plot and subset to check data
# WS1data %>%
#   filter(DATE > mdy("4-01-1986") & DATE < mdy("11-01-1986")) %>%
#   ggplot(aes(DATE, Q, color = "daily"))+
#     geom_line()+
#     geom_line(aes(x = DATE, y = Low7Day, color = "7 day mean"))

# testSubset<-WS2data %>%
#   filter(DATE > mdy("01-01-1987") & DATE < mdy("12-31-1987"))

#Create annual low flow, along with total precip and date of low flow, add precip for preceding WY
#Discard 1979 and 2018 for incomplete data
WS1LowFlow <- WS1data %>% 
  mutate(Year = year(DATE)) %>% 
  group_by(Year) %>% 
  summarize(minQ = min(Low7Day, na.rm = TRUE),
            minQDate = DATE[which.min(Low7Day)],
            annP = sum(precip, na.rm = TRUE)) %>% 
  mutate(minQDoy = yday(minQDate))

WS1WYprecip<- WS1data %>% 
  group_by(WY) %>% 
  summarize(annPWY = sum(precip, na.rm = TRUE))

WS1LowFlow$annPWY <- WS1WYprecip$annPWY

WS1LowFlow <- WS1LowFlow %>% 
  filter(Year != 1979 & Year != 2018)

#Some figures summarizing this annual dataset
WS1LowFlow %>% 
  ggplot(aes(y = minQ)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min daily runoff (mm)")
WS1LowFlow %>% 
  ggplot(aes(x = Year, y = minQ)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min daily runoff (mm)")

WS1LowFlow %>% 
  ggplot(aes(y = minQDoy)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "7 day min day of year (mm)") 
WS1LowFlow %>% 
  ggplot(aes(x = Year, y = minQDoy)) +
  geom_bar(stat = 'identity') +
  labs(y = "7 day min day of year (mm)")

WS1LowFlow %>% 
  ggplot(aes(y = annPWY)) +
  geom_boxplot() +
  scale_x_discrete() +
  labs(y = "annual precip (mm)")
WS1LowFlow %>% 
  ggplot(aes(x = Year, y = annPWY)) +
  geom_bar(stat = 'identity') +
  labs(y = "annual precip (mm)")
  
#Figures comparing them
# ggplot(data = WS2LowFlow, aes(x = annP, y = minQ))+
#   geom_point()+
#    labs(x = "annual precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS1LowFlow, aes(x = annPWY, y = minQ))+
  geom_point()+
   labs(x = "prior WY precip (mm)", y = "7 day min daily runoff (mm)")
ggplot(data = WS1LowFlow, aes(x = minQDoy, y = minQ))+
  geom_point()+
  labs(x = "7 day min day of year (mm)", y = "7 day min daily runoff (mm)")
```


Exploratory analysis using WSXLowFlow data and climate variables with climwin package. Separate execution bc of runtime
- extract data for WS2 in this case
- slidingwin does not work with NAs in the predictive (climate) data. Replace those values with 0 for now
```{r ClimaWinData}
climData <- allData %>% 
  select(DATE,precip) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y"))

climData$precip[is.na(climData$precip)] <- 0

lowFlowData <- WS2LowFlow %>% 
  select(minQDate,minQ) %>% 
  mutate(txtData = format(minQDate,"%d/%m/%Y"))
```

Execute climwin main function 
```{r ClimaWinDataExec}
winOut <- slidingwin(xvar = list(climData$precip),
                     cdate = climData$txtDate,
                     bdate = lowFlowData$txtData,
                     baseline = lm(minQ ~ 1, data = lowFlowData),
                     cinterval = "day",
                     range = c(365,0),
                     type = "relative",
                     stat = "mean",
                     func = "lin"
                     )

winModels <- winOut[[1]]$Dataset
```

Visualize data tables and plots using built in functions
```{r ClimaWinDataVis}
head(winModels)

plotdelta(dataset = winModels)
plotweights(dataset = winModels)

```

```

