---
title: "Low Flow Figures"
output:
  html_document:
    df_print: paged
---
Load packages
```{r packages, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(climwin)
library(plotly)
library(imputeTS)
library(ggpubr)
library(ggthemes)
library(modelr)
library(tidyverse)
```

Load data
  - Transpiration deficit data
```{r loadData, echo=FALSE}
# load("../TransDef.RData") #If being knit
load("TransDef.RData")
WS1Profile <- read.csv("Data/WS01Profile.csv")
```


Figure 1: Site map

Figure 2: Hydro and climate data

Figure 3: Hypothesis figures
```{r HypothesisFigures}
WS1Plot <- WS1Profile %>% 
  mutate(SoilDepth = Elev_m - 2) %>% 
  ggplot() +
  geom_line(aes(x = Dist_m, y = Elev_m)) +
  geom_line(aes(x = Dist_m, y = SoilDepth), color = "brown") +
  theme_bw()
  # theme(axis.text.x = element_blank(), axis.text.y = element_blank()) +
  # labs(x = "", y = "")

print(WS1Plot)
ggsave("ManuscriptDocuments/Figures/WS1Profile.pdf", plot = WS1Plot)
```


Figure 4: Window analysis

Figure 5: Transpiration deficit
```{r TransDefFig}
youngCol = "#4dac26"
oldCol = "#a6611a"
ripDefCol = "#01665e"
hillDefCol = "#80cdc1"

myCols = c(c("Young" = youngCol, "Old" = oldCol, "HillDef" = hillDefCol, "RipDef" = ripDefCol))

myLineWidth = 2
myTxtSz = 20

#Hillslope Transpiration Deficit
HillslopeTransDefPlot <- TransDef %>% 
  select(DATE, transHillOld, transHillYoung, transDefHill) %>% 
  ggplot() +
  geom_area(aes(x = DATE, y = transHillYoung, fill = "Young"), 
            alpha = 0.3) + 
  geom_area(aes(x = DATE, y = transHillOld, fill = "Old"), 
            alpha = 0.3) +
  geom_area(aes(x = DATE, y = transDefHill, fill = "HillDef"), 
            alpha = 0.7) +
  theme_bw(base_size = myTxtSz) +
  scale_x_date(expand = c(0,0)) +
  labs(x = "", y = "Transpiration (mm/d)") +
  scale_fill_manual(name = "", values = myCols)
print(HillslopeTransDefPlot)

#Riparian Transpiration Deficit
RiparianTransDefPlot <- TransDef %>% 
  select(DATE, transRipOld, transRipYoung, transDefRip) %>% 
  ggplot() +
  geom_area(aes(x = DATE, y = transRipYoung, fill = "Young"), 
            alpha = 0.3) + 
  geom_area(aes(x = DATE, y = transRipOld, fill = "Old"), 
            alpha = 0.3) +
  geom_area(aes(x = DATE, y = transDefRip, fill = "RipDef"),
            alpha = 0.7) +
  theme_bw(base_size = myTxtSz) +
  scale_x_date(expand = c(0,0)) +
  labs(x = "", y = "Transpiration (mm/d)") +
  scale_fill_manual(name = "", values = myCols)
print(RiparianTransDefPlot)

#Overall Transpiration Deficit
myCols = c(Rip = ripDefCol, Hill = hillDefCol)

OverallTransDefPlot <- TransDef %>% 
  select(DATE, transDefRip, transDefHill) %>% 
  mutate(Rip = cumsum(transDefRip),
         Hill = cumsum(transDefHill)) %>% 
  select(-c(transDefRip, transDefHill)) %>% 
  gather(Position, sumTrans, Rip, Hill) %>% 
  ggplot(aes(x = DATE, y = sumTrans, fill = Position)) +
  geom_area(alpha = 0.7) +
  scale_fill_manual(name = "", values = myCols) +
   theme_bw(base_size = myTxtSz) +
  scale_x_date(expand = c(0,0)) +
  labs(x = "", y = "\u03A3 Transpiration (mm)")
print(OverallTransDefPlot)  

```


S1: Flow deficit

S2: Digitized Moore data

S3: Regressions b/n sap flux density and radiation/vpd

S4: Riparian well time series

S5: Hillslope well time series or soil volumetric content + porosity

S6: Table of calculation values