---
title: "HJAClimWin"
output: html_notebook
---

This file applies the base function in the climwin package to examine relationships between climate variables (precip and VPD) at HJA and late summer baseflow
in WS1, WS2, WS3, WS9, WS10. Then it restructures the output data and visualizes them using custom functions instead of the climwin baked in visualizations. 

Load packages
```{r LoadPackages}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(climwin)
library(plotly)
library(imputeTS)
library(ggpubr)
library(ggthemes)
library(modelr)
```


Get precip data from allData.
- slidingwin does not work with NAs in the predictive (climate) data. Replace those values with 0

```{r PrecipData}
climDataPrecip <- allData %>% 
  select(DATE, precip, meanTemp, meanVpd) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y")) %>% #Format data for use with slidingwin function
  filter(DATE > as.Date("1998-01-01")) #For when v-notch weirs were put on watersheds revealing error in WS1

climDataPrecip$precip[is.na(climDataPrecip$precip)] <- 0
```

Get daytime VPD data from that provided by KJ

```{r KJdata}
load("Data/EDI_met_daily_daytime_nighttime_stats_20220505.Rdat")

# ggplot_na_distribution(metdaytime_final$vpdmean, x_axis_labels = metdaytime_final$date)

climDataVpd <- metdaytime_final %>% 
  select(date, vpdmean) %>% 
  rename(DATE = date, meanVpd = vpdmean) %>% 
  mutate(txtDate = format(DATE,"%d/%m/%Y")) %>% 
  filter(year(DATE) >= 1998) #When weirs were put on WSs
```

Get low flow data for WSXLowFlow data
- extract data for all WS low flow Aug 30

```{r LowFlowData}
lowFlowData <- WSLowFlowAnn %>% 
  ungroup() %>% 
  select(WS1minQAug, WS2minQAug, WS3minQAug, WS9minQAug, WS10minQAug, Year) %>% 
  mutate(txtDate = paste0("30/08/",Year)) %>% 
  filter(Year > 1998) #V-notch weirs see above note
```

Executing sliding window for both precip and vpd and save default output as a single structure.
- Output is saved at end of chunk because runtime is ~5 min on 2.6 GHz 6 core Intel i7
```{r ClimaWinDataExec2}
# Select which watersheds to use out of all possible:
# WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug", "WS9minQAug", "WS10minQAug")
WSvarnames <- c("WS1minQAug", "WS2minQAug")

#preallocate lists
climWinModelsPrecip <- vector(mode = "list", length = length(WSvarnames)) 
climWinModelsVpd <- vector(mode = "list", length = length(WSvarnames)) 

for (i in 1:length(WSvarnames)){
  lowFlowTemp <- lowFlowData %>% #Extract specific WS and remove rows with NA
    select(Year, txtDate, minQ = WSvarnames[i]) %>% 
    drop_na()
  
  #Precip  
  winOut <- slidingwin(xvar = list(climDataPrecip$precip),
                     cdate = climDataPrecip$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,10),
                     type = "relative",
                     stat = "sum",
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

  climWinModelsPrecip[[i]] <- winOut[[1]]$Dataset
  
  #Vpd
  winOut <- slidingwin(xvar = list(climDataVpd$meanVpd),
                     cdate = climDataVpd$txtDate,
                     bdate = lowFlowTemp$txtDate,
                     baseline = lm(minQ ~ 1, data = lowFlowTemp),
                     cinterval = "day",
                     range = c(365,10),
                     type = "relative",
                     stat = "mean",
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

  climWinModelsVpd[[i]] <- winOut[[1]]$Dataset
}

#Save these files
save(climWinModelsPrecip, climWinModelsVpd, file = "climWinOutputs.RDat")
```

Filter models to only those greater than some length of time and look at box plots and tile plots of top X open and close windows. Visualize these results with boxplots and tile plots.
```{r TopModels}
#inputs
winLength <- 14 #number of days of minimum window
numTopModels <- 40 #top number of models as ranked by AIC to plot

#Pre-allocate data frame
topClimModels <- data.frame(deltaAICc = double(),
                            WindowOpen = integer(),
                            WindowClose = integer(),
                            WSname = character(),
                            climPred = character())

for (i in 1:length(WSvarnames)) {
  #Precip models
  tempClimModels <- climWinModelsPrecip[[i]] %>% 
    select(deltaAICc, WindowOpen, WindowClose) %>% 
    mutate(WSname = str_sub(WSvarnames[i], 1, -8),
           climPred = "precip") %>% 
    mutate(lengthWin = WindowOpen - WindowClose + 1) %>% 
    filter(lengthWin >= winLength) %>% 
    select(-lengthWin) %>% 
    slice_min(deltaAICc, n = numTopModels)
  
  topClimModels <- rbind(topClimModels,tempClimModels)
  
  #Vpd models
  tempClimModels <- climWinModelsVpd[[i]] %>% 
    select(deltaAICc, WindowOpen, WindowClose) %>% 
    mutate(WSname = str_sub(WSvarnames[i], 1, -8),
           climPred = "vpd") %>% 
    mutate(lengthWin = WindowOpen - WindowClose + 1) %>% 
    filter(lengthWin >= winLength) %>% 
    select(-lengthWin) %>% 
    slice_min(deltaAICc, n = numTopModels)
  
  topClimModels <- rbind(topClimModels,tempClimModels)
} 

#Plot tile plots of top models
winDaysVec <- min(topClimModels$WindowClose):max(topClimModels$WindowOpen) #Create vector of all days in any window 

WSPredCombos <- unique(topClimModels[c("WSname", "climPred")])
WSCombos <- WSPredCombos$WSname
PredCombos <- WSPredCombos$climPred

tileTopModels <- data.frame(winDay = rep(winDaysVec, each = nrow(WSPredCombos))) %>%
  mutate(WS = rep(WSCombos, times = length(winDaysVec))) %>% 
  mutate(climVar = rep(PredCombos, times = length(winDaysVec))) %>% 
  mutate(nModDay = 0)

for (i in 1:nrow(tileTopModels)) {
  tempTopClimModels <- topClimModels %>% 
    filter(WSname == tileTopModels$WS[i] & 
             climPred == tileTopModels$climVar[i]) %>% 
    filter(WindowClose <= tileTopModels$winDay[i] &
             WindowOpen >= tileTopModels$winDay[i])
  
  tileTopModels$nModDay[i] <- nrow(tempTopClimModels)
}

tileTopModels$winDay <- winDays2Dates(tileTopModels$winDay) #Convert window days into actual dates

save(tileTopModels, file = "ManuscriptDocuments/Figures/FigureData/ClimWinTile.RData")

tilePlots <- tileTopModels %>% 
  ggplot(aes(x = winDay, y = WS, fill = nModDay)) +
  geom_tile() + facet_wrap(~ climVar) +
  labs( # title = paste("Top ", numTopModels, " models explaining Aug 30 low flow"),
       x = "Open and close days",
       y = "Watershed",
       fill = "# of models"
       ) +
  theme_bw(base_size = 20)
print(tilePlots)
ggplotly(tilePlots)
```

Function is for converting window open and window close days to dates for plotting purposes
```{r days2datesBeforeAug30}
winDays2Dates <- function(d) {
  doys <- 241 - d #Convert day before Aug 30 to day of year. R starts from day 0 not 1
  output <- as.Date(doys, origin = "2023-01-01")
}
```

NOT USED FOR MANUSCRIPT

Look at individual model fits. Choose which model to plot based on best models output from slidingwin output and then use singlewin to get the specific model
```{r plotIndvModels}
winOpen <- 356
winClose <- 1
ws <- 2 #Which watershed: WS1, WS2, WS3, WS9, WS10
climVar <- "Precip" #Precip or VPD

WSvarnames <- c("WS1minQAug","WS2minQAug","WS3minQAug", "WS9minQAug", "WS10minQAug")
WSnames <- str_sub(WSvarnames, 1, -8)

#Select correct climate data
if (climVar == "Precip"){
  climDataTemp <- climDataPrecip %>% 
    select(txtDate, climPredictor = precip)
} else if (climVar == "VPD"){
  climDataTemp <- climDataVpd %>% 
    select(txtDate, climPredictor = meanVpd)
} else {
  stop("climVar must be precip vpd")
}

#Select correct watershed data
lowFlowDataTemp <- lowFlowData %>% 
  select(txtDate, minQAug = WSvarnames[ws])

#Select correct method for if it's precip (sum) or vpd (mean)
if (climVar == "Precip") {
  statType <- "sum"
} else if (climVar == "VPD") {
  statType <- "mean" 
} else {
  stop("climVar must be precip vpd")
}

#Get model to plot
model2plot <- singlewin(xvar = list(climDataTemp$climPredictor),
                     cdate = climDataTemp$txtDate,
                     bdate = lowFlowDataTemp$txtDate,
                     baseline = lm(minQAug ~ 1, data = lowFlowDataTemp),
                     cinterval = "day",
                     range = c(winOpen, winClose),
                     type = "relative",
                     stat = statType,
                     func = "lin",
                     cmissing = "method2" #This method replaces missing values with mean value of all records with same date
                     )

#Select correct output dataset for climate variable
if (climVar == "Precip") {
  climDataset <- climWinModelsPrecip[[ws]]
} else if (climVar == "VPD") {
  climDataset <- climWinModelsVpd[[ws]]
} else {
  stop("climVar must be 1 for precip or 2 for vpd")
}
  
plotbest(dataset = climDataset, 
         bestmodel = model2plot$BestModel,
         bestmodeldata = model2plot$BestModelData)

#Manual version of above plot
modelPlot <- ggplot(data = model2plot$BestModelData, aes(x = climate, y = yvar)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() + 
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label") +
  labs(title = paste("Window", climVar, "vs Aug 30 LF for", WSnames[ws]),
       y = "7 day mean Q (mm/day)",
       x = "Total precip (mm)")
print(modelPlot)
```

When analysis was run using only data from post v-notch weirs, WS1 had a strong set of models in the first to second week of september. Exploring what might have caused that and whether it appears to be spurious or not. A single extreme value is levering R2
```{r SeptWinInWS1}
#Look at low flow values for each WS
lowFlowPlot <- lowFlowData %>% 
  select(Year,
         WS1 = WS1minQAug, 
         WS2 = WS2minQAug,
         WS3 = WS3minQAug) %>% 
  gather(key = WS, value = LF, WS1:WS3) %>% 
  ggplot(aes(x = Year, y = LF, color = WS)) + geom_line()

print(lowFlowPlot)

#Look at some of those models:
# Highest model for WS1
plot_indv_win_mod(358, 356, 1, "Precip")
# Same window for WS3
plot_indv_win_mod(358, 356, 3, "Precip")

# Models closer to original window
plot_indv_win_mod(95, 83, 1, "Precip")
plot_indv_win_mod(95, 83, 3, "Precip")

```

Function replicating content of PlotIndvModels chunk
```{r plotIndvWinModelsFunction}
#Inputs:
# WinO and WinC are open and close days from climwin outputs
# wsNum is which of wss 1-3 (1-3), 9 (4), 10 (5)
# cVar is either "Precip" or "VPD"

#Requires in Environment:
# climDataPrecip and climDataVpd
# lowFlowData
# climWinModelsPrecip
plot_indv_win_mod <- function(winO, winC, wsNum, cVar) {
  
  WSLFnames <- c("WS1minQAug","WS2minQAug","WS3minQAug", "WS9minQAug", "WS10minQAug")
  WS <- str_sub(WSvarnames, 1, -8)
  
  #Select correct climate data
  if (cVar == "Precip"){
    climDataFnc <- climDataPrecip %>% 
      select(txtDate, climPredictor = precip)
  } else if (cVar == "VPD"){
    climDataFnc <- climDataVpd %>% 
      select(txtDate, climPredictor = meanVpd)
  } else {
    stop("cVar must be precip vpd")
  }
  
  #Select correct watershed data
  lowFlowDataFnc <- lowFlowData %>% 
    select(txtDate, minQAug = WSLFnames[wsNum])
  
  #Select correct method for if it's precip (sum) or vpd (mean)
  if (cVar == "Precip") {
    statType <- "sum"
  } else if (cVar == "VPD") {
    statType <- "mean" 
  } else {
    stop("cVar must be precip vpd")
  }
  
  #Get model to plot
  model2plotFnc <- singlewin(xvar = list(climDataFnc$climPredictor),
                          cdate = climDataFnc$txtDate,
                          bdate = lowFlowDataFnc$txtDate,
                          baseline = lm(minQAug ~ 1, data = lowFlowDataFnc),
                          cinterval = "day",
                          range = c(winO, winC),
                          type = "relative",
                          stat = statType,
                          func = "lin",
                          cmissing = "method2" #This method replaces missing values with mean value of all records with same date
  )
  
  #Manual version of above plot
  modelPlot <- ggplot(data = model2plotFnc$BestModelData, aes(x = climate, y = yvar)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_bw() + 
    stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label") +
    labs(title = paste("Window", cVar, "vs Aug 30 LF for", WS[wsNum],
                       format(winDays2Dates(winO), format = "%m-%d"), "-",
                       format(winDays2Dates(winC), format = "%m-%d")),
         y = "7 day mean Q (mm/day)",
         x = "Total precip (mm)")
  print(modelPlot)
}
```