---
title: "HJA_ET_fromMoore2004"
output:
  html_document:
    df_print: paged
---

Load packages
```{r LoadPackages}
library(lubridate)
library(ggplot2)
library(zoo)
library(climwin)
library(plotly)
library(imputeTS)
library(ggpubr)
library(ggthemes)
library(modelr)
library(tidyverse)
```
Attempts to leverage data from Moore et al 2004 study on transpiration at HJA WS1 and 2

Load data from digitized csvs
```{r Data}
Figure3aData <- read.csv("Data/Moore2004Data/Figure3a.csv")
Figure3bData <- read.csv("Data/Moore2004Data/Figure3b.csv")
Figure3cData <- read.csv("Data/Moore2004Data/Figure3c.csv")
Figure7Data <- read.csv("Data/Moore2004Data/Figure7.csv")

```

Restructure data:
- Make three df (sap flux 1999 and 2000 and transpiration)
- Lots of fiddling to get day of year into the right date
- Aggregate into daily values
- Linear interpolation of missing values
- Figures 3a and 3c have PSMEo data, but 3c over a longer duration, so use that data in 2000 sapFlux

```{r structureData}
sapFlux1999 <- Figure3bData %>% 
  rename(sFlux = SapFluxDens) %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, tree) %>% 
  summarise(sFluxDaily = mean(sFlux)) %>% 
  ungroup() %>% 
  spread(tree,sFluxDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(PSMEy = na.approx(PSMEy),
         ALRUy = na.approx(ALRUy)) %>% 
  gather(tree, sFlux, PSMEy, ALRUy) %>% 
  filter(!between(DATE, as.Date("1999-07-29"), as.Date("1999-08-04"))) #Data gap in figure 3b. Interpolate later

sapFlux20003a <- Figure3aData %>% 
  rename(sFlux = SapFluxDens) %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, tree) %>% 
  summarise(sFluxDaily = mean(sFlux)) %>% 
  ungroup() %>% 
  spread(tree,sFluxDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(PSMEy = na.approx(PSMEy),
         PSMEo = na.approx(PSMEo)) %>% 
  gather(tree, sFlux, PSMEy, PSMEo) %>% 
  filter(tree == "PSMEy") #Since PSMEo is in Figure 3c data don't need it here

sapFlux2000 <- Figure3cData %>% 
  rename(sFlux = SapFluxDens) %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, tree) %>% 
  summarise(sFluxDaily = mean(sFlux)) %>% 
  ungroup() %>% 
  spread(tree,sFluxDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(TSHEo = na.approx(TSHEo, na.rm = FALSE, rule = 2),
         PSMEo = na.approx(PSMEo)) %>% 
  gather(tree, sFlux, TSHEo, PSMEo) %>% 
  bind_rows(sapFlux20003a) #Add PSMEy from figure 3a
  
trans2000 <- Figure7Data %>% 
  mutate(daysInYear = ifelse(leap_year(year), 366, 365)) %>% 
  mutate(decimalDate = year + DOY/daysInYear) %>% 
  mutate(dateTime = date_decimal(decimalDate)) %>% 
  mutate(DATE = as.Date(dateTime)) %>% 
  group_by(DATE, stand) %>% 
  summarise(transDaily = mean(trans)) %>% 
  ungroup() %>% 
  spread(stand,transDaily) %>% 
  padr::pad(interval = "day") %>% 
  mutate(young = na.approx(young),
         old = na.approx(old)) %>% 
  gather(stand, trans, young, old)

#Plot these dfs to check how they match with Moore 2004
Figure3aMoore <- sapFlux2000 %>% 
  ggplot(aes(x = DATE, y = sFlux)) +
  geom_line(aes(color = tree))
print(Figure3aMoore)

Figure3bMoore <- sapFlux1999 %>%
ggplot(aes(x = DATE, y = sFlux)) +
  geom_line(aes(color = tree))
print(Figure3bMoore)

Figure7Moore <- trans2000 %>% 
  ggplot(aes(x = DATE, y = trans)) +
  geom_line(aes(color = stand))
print(Figure7Moore)

```

Now get vpd and radiation data for these same time periods but also as early as march or so to back extrapolate
```{r metData4Trans}
vpdMoore <- allData %>% 
  filter(between(DATE, as.Date("1999-03-01"), as.Date("2000-11-01"))) %>% 
  select(DATE,meanVpd) %>% 
  ggplot(aes(x = DATE, y = meanVpd)) + geom_line()

print(vpdMoore)

#Missing data from 2000-6-4 - 2000-6-21 and 2000-5-24 - 2000-5-31. Check if there are other data in this range at primet
vpdCheck <- vpdRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE,VPD_MEAN_DAY,SITECODE,HEIGHT,PROBE_CODE) %>% 
  filter(between(DATE, as.Date("2000-05-23"), as.Date("2000-06-01")) |
         between(DATE, as.Date("2000-06-03"), as.Date("2000-06-22"))) %>% 
  filter(!is.na(VPD_MEAN_DAY)) %>% 
  filter(SITECODE %in% c("PRIMET", "CS2MET"))

#Find radiation data for this period
radRaw <- read.csv("Data/MS00105_v6.csv")

metRad <- radRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE, SOLAR_TOT_DAY, SITECODE, HEIGHT, PROBE_CODE) %>% 
  filter(between(DATE, as.Date("1999-03-01"), as.Date("2000-11-01"))) %>% 
  filter(SITECODE == "PRIMET")

radPlot <- metRad %>% 
  ggplot(aes(x = DATE, y = SOLAR_TOT_DAY)) + geom_line()

print(radPlot)
#Data gap at 2000-5-24 - 2000-5-31
```

Check to see what other sensors and stations could be suitable for linear regression interpolation.
  Requires radRaw loaded in previous chunk from MS00105_v6.csv.
  First select a data range and create vpd and rad dfs
```{r gapFillingVpdRad1}
fromToDates <- c("1999-03-01", "2000-11-01")

#Vpd data from all sites, sensors, heights
vpdCheck <- vpdRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE,VPD_MEAN_DAY,SITECODE,HEIGHT,PROBE_CODE) %>% 
  filter(between(DATE, as.Date(fromToDates[1]), as.Date(fromToDates[2])))

#Radiation data from all sites, sensors, heights
radCheck <- radRaw %>% 
  mutate(DATE = ymd(DATE)) %>% 
  select(DATE,SOLAR_TOT_DAY,SITECODE,HEIGHT,PROBE_CODE) %>% 
  filter(between(DATE, as.Date(fromToDates[1]), as.Date(fromToDates[2])))
```

  Then get unique combos of site and height and test relationships between each of these and primet data (probe 04, not 05) 
```{r gapFillingVpdRad2}
varCheck <- vpdCheck
datesGap <- c("2000-05-25", "2000-05-30", "2000-06-05", "2000-06-20") #gaps for VPD
# datesGap <- c("2000-05-25", "2000-05-30") #gaps for radiation

uniqueCombos <- unique(varCheck[ , c("SITECODE","HEIGHT")]) %>% 
  filter(SITECODE != "PRIMET") %>% 
  mutate(numNAs = 0)

for (i in 1:nrow(uniqueCombos)) { #:nrow(uniqueCombos)
  varPRI <- varCheck %>% 
    filter(SITECODE == "PRIMET") %>% 
    filter(PROBE_CODE == "VPDPRI04" |
             PROBE_CODE == "RADPRI01")
  varOther <- varCheck %>% 
    filter(SITECODE == uniqueCombos$SITECODE[i]) %>% 
    filter(HEIGHT == uniqueCombos$HEIGHT[i])
  
  #Check how many NAs are in potential predictor data
  varOtherCheck <- varOther %>% 
    filter(between(DATE, as.Date(datesGap[1]), as.Date(datesGap[2])) |
           between(DATE, as.Date(datesGap[3]), as.Date(datesGap[4])))
  
  uniqueCombos$numNAs[i] <- sum(is.na(varOtherCheck$VPD_MEAN_DAY))
  if(uniqueCombos$numNAs[i] != 0) next
  
  #Plot bivariate of predictor and PRIMET data
  tempPlot <- left_join(varPRI, varOther, by = "DATE") %>% 
    select(DATE, 2, 6) %>% 
    rename(PRIvar = 2, OtherVar = 3) %>% 
    drop_na() %>% 
    ggplot(aes(x = OtherVar, y = PRIvar)) + geom_point() +
    geom_smooth(formula = y ~ x, method = "lm") +
    stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label") +
    labs(title = paste(uniqueCombos$SITECODE[i], uniqueCombos$HEIGHT[i], "cm"))
  print(tempPlot)
}
```

Finally use the best models to fill gaps in PRIMET data and save as a single data frame
  Best Models:
  - VPD: CS2MET, 150 cm
  - Radiation: CENMET, 625 cm
  Requires vpdCheck and radCheck from two chunks up
```{r gapFillingVpdRad3}
#Put PRIMET radiation, vpd in same df as predictor variables
PRIgfVpd <- vpdCheck %>% 
  filter(SITECODE %in% c("PRIMET", "CS2MET")) %>% 
  filter(PROBE_CODE != "VPDPRI05") %>% 
  select(DATE, VPD_MEAN_DAY, SITECODE) %>% 
  spread(SITECODE, VPD_MEAN_DAY) 

PRIgfRad <- radCheck %>% 
  filter(SITECODE %in% c("PRIMET", "CENMET")) %>% 
  select(DATE, SOLAR_TOT_DAY, SITECODE) %>% 
  spread(SITECODE, SOLAR_TOT_DAY) 

#Find lm, add prediction column, replace NAs in PRIMET with prediction
vpdReg <- lm(PRIMET ~ CS2MET, data = PRIgfVpd)

PRIgfVpd <- PRIgfVpd %>% 
  add_predictions(vpdReg) %>% 
  mutate(gf = ifelse(is.na(PRIMET), 1, 0)) %>% 
  mutate(PRIMET = ifelse(is.na(PRIMET), pred, PRIMET)) %>% 
  select(DATE, PRIMET, gf) %>% 
  mutate(PRIMET = na.approx(PRIMET))

radReg <- lm(PRIMET ~ CENMET, data = PRIgfRad)

PRIgfRad <- PRIgfRad %>% 
  add_predictions(radReg) %>% 
  mutate(gf = ifelse(is.na(PRIMET), 1, 0)) %>% 
  mutate(PRIMET = ifelse(is.na(PRIMET), pred, PRIMET)) %>% 
  select(DATE, PRIMET, gf) %>% 
  mutate(PRIMET = na.approx(PRIMET))

#Plot data highlighting gap-fill
vpdGfPlot <- PRIgfVpd %>% 
  ggplot(aes(x = DATE, y = PRIMET)) + geom_line(aes(color = gf))
print(vpdGfPlot)
  
radGfPlot <- PRIgfRad %>% 
  ggplot(aes(x = DATE, y = PRIMET)) + geom_line(aes(color = gf))
print(radGfPlot)

#Combine gap-filled vpd and radiation data into single df
gfPRI <- left_join(PRIgfRad, PRIgfVpd, by = "DATE") %>% 
  select(DATE, Rad = PRIMET.x, Vpd = PRIMET.y)
```

Now, use the gap filled vpd and radiation data to explore extrapolating transpiration data reported in Moore et al.
  Using:
  - trans2000
  - sapFlux1999
  - sapFlux2000
  - gfPRI
```{r ExtrapolateTrans}
transEx <- trans2000 %>% 
  spread(stand, trans) %>% 
  right_join(., gfPRI, by = "DATE")

#Plot bivariates of transpiration against radiation and against vpd
# transRadPlot <- transEx %>%
#   select(DATE, old, young, Rad) %>%
#   drop_na() %>%
#   ggplot() +
#   geom_point(aes(x = Rad, y = old, color = "old")) +
#   geom_point(aes(x = Rad, y = young, color = "young")) +
#   labs(y = "Trans")
# print(transRadPlot)
# 
# transVpdPlot <- transEx %>%
#   select(DATE, old, young, Vpd) %>%
#   drop_na() %>%
#   ggplot() +
#   geom_point(aes(x = Vpd, y = old, color = "old")) +
#   geom_point(aes(x = Vpd, y = young, color = "young")) +
#   labs(y = "Trans")
# print(transVpdPlot)

# Generate a bunch of regression models. 
RegTransOldVpd <- lm(old ~ Vpd, data = transEx)
RegTransOldRad <- lm(old ~ Rad, data = transEx)
RegTransOldVpdRad <- lm(old ~ Rad + Vpd, data = transEx)
RegTransYoungVpd <- lm(young ~ Vpd, data = transEx)
RegTransYoungRad <- lm(young ~ Rad, data = transEx)
RegTransYoungVpdRad <- lm(young ~ Rad + Vpd, data = transEx)

summary(RegTransOldVpd)
summary(RegTransOldRad)
summary(RegTransOldVpdRad)
summary(RegTransYoungVpd)
summary(RegTransYoungRad)
summary(RegTransYoungVpdRad)
# Rad is best predictor by a little. MLM not much of an improvement over just Rad

#Use Radiation to back predict data through April 1
transExPred <- transEx %>% 
  add_predictions(model = RegTransOldRad, var = "oldPred") %>% 
  add_predictions(model = RegTransYoungVpd, var = "youngPred") %>% 
  mutate(Extrap = ifelse(is.na(old),1,0)) %>% 
  mutate(old = ifelse(is.na(old), oldPred, old)) %>% 
  mutate(young = ifelse(is.na(young), youngPred, young)) 

transPredYoungPlot <- transExPred %>%  
  filter(between(DATE, as.Date("2000-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = young, color = Extrap))
print(transPredYoungPlot)

transPredOldPlot <- transExPred %>%  
  filter(between(DATE, as.Date("2000-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = old, color = Extrap))
print(transPredOldPlot)

radPlot <- transExPred %>% 
  filter(between(DATE, as.Date("2000-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = Rad))
print(radPlot)
```
Check to see if I can go from sap flux measurements to transpiration figure
- Input SBA values from fig 6c,d
- Calculate a 2000 Alder sap flux series from eqn 4 relationship
- Apply eqn 2 (and divide by 10e4 to account for unit conversions)
```{r CheckFig3ToFig7}
Tree <- c("PSME", "TSHE", "ALRU", "Ang", "Gym")
Old <- c(5.5, 9.9, 0.2, 1.3, 15.9)
Young <- c(13.9, 0.1, 1.5, 7.7, 13.9)

SBAFig6 <- data.frame(Tree, Old, Young) 

sapFlux2000check <- sapFlux2000 %>% 
  spread(tree, sFlux) %>%
  mutate(doy = yday(DATE)) %>% 
  mutate(ALRUy = PSMEy * (-163 + 165.1 * (1 - exp(-0.0293 * doy)))) %>% 
  select(-"doy") %>% 
  gather(key = tree, value = sFlux, PSMEo, PSMEy, TSHEo, ALRUy)

sapFluxPlot <- sapFlux2000check %>% 
  ggplot(aes(x = DATE, y = sFlux, color = tree)) + geom_line()
print(sapFluxPlot)

trans2000check <- sapFlux2000check %>% 
  spread(tree, sFlux) %>% 
  mutate(TransYoung = (PSMEy*13.9 + ALRUy*7.7)/10000) %>% 
  mutate(TransOld = (PSMEo*5.5 + TSHEo*9.9 + ALRUy*1.3)/10000)

transCheckPlot <- trans2000check %>% 
  select(DATE, TransYoung, TransOld) %>% 
  gather(Stand, Trans, TransYoung, TransOld) %>% 
  ggplot(aes(x = DATE, y = Trans, color = Stand)) + geom_line()

ggplotly(transCheckPlot)
```

Get regression relationship between radiation/vpd and 1999 ALRU sapflux, use to extrapolate to 2000 ALRU sapflux.
  - Uses sapFlux1999 and gfPRI
  - Met data only during extrapolation period (April 1 - Nov 1)
```{r ExtrapolateALRUSapFlux}
ALRUEx <- sapFlux1999 %>% 
  filter(tree == "ALRUy") %>% 
  select(DATE, sFlux) %>% 
  right_join(gfPRI, by = "DATE")

#Plot bivariates of each
ALRUvRadPlot <- ALRUEx %>% 
  drop_na() %>% 
  ggplot(aes(x = Rad, y = sFlux)) + 
  geom_point() + 
  geom_smooth(formula = y ~ x, method = "lm") +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
print(ALRUvRadPlot)

ALRUvVpdPlot <- ALRUEx %>% 
  drop_na() %>% 
  ggplot(aes(x = Vpd, y = sFlux)) + 
  geom_point() + 
  geom_smooth(formula = y ~ x, method = "lm") +
  stat_cor(aes(label = after_stat(rr.label)), color = "red", geom = "label")
print(ALRUvVpdPlot)

#Generate linear models and check them
RegALRUVpd <- lm(sFlux ~ Vpd, data = ALRUEx)
RegALRURad <- lm(sFlux ~ Rad, data = ALRUEx)
RegALRUVpdRad <- lm(sFlux ~ Rad + Vpd, data = ALRUEx)
RegALRUVpdXRad <- lm(sFlux ~ Rad*Vpd, data = ALRUEx)

summary(RegALRUVpd)
summary(RegALRURad)
summary(RegALRUVpdRad)
summary(RegALRUVpdXRad)

#Use vpd and rad to extrapolate ALRU to April 1 - Nov 1 2000
ALRUExPred <- ALRUEx %>% 
  add_predictions(model = RegALRUVpdRad, var = "sFluxPred") %>% 
  mutate(Extrap = ifelse(is.na(sFlux), 1, 0)) %>% #To mark what data is extrapolated vs measured
  mutate(sFlux = ifelse(is.na(sFlux), sFluxPred, sFlux))
  
#Plot extrapolated data
ALRUPredPlot <- ALRUExPred %>% 
  filter(between(DATE, as.Date("1999-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = sFlux, color = Extrap))
print(ALRUPredPlot)
```

Do the same for TSHEo, PSMEo, and PSMEy: Find best linear models
  - PSMEo has lowest R2
  - Best models generally either vpd or both vpd and rad
  - Since above (ALRUy) uses both predictors do the same here
```{r ExtrapolateSapFlux}
sapFluxEx <- sapFlux2000 %>% 
  spread(tree, sFlux) %>% 
  right_join(gfPRI, by = "DATE")

#Plot bivariates between rad/vpd and each tree species/age
#Radiation
# for (i in 2:4) {
#   tempRadPlot <- sapFluxEx %>% 
#     select(Rad, i) %>% 
#     drop_na() %>% 
#     ggplot(aes_string(colnames(.)[1], colnames(.)[2])) + geom_point()
#   print(tempRadPlot)
# }
# 
# #Vpd
# for (i in 2:4) {
#   tempVpdPlot <- sapFluxEx %>% 
#     select(Vpd, i) %>% 
#     drop_na() %>% 
#     ggplot(aes_string(colnames(.)[1], colnames(.)[2])) + geom_point()
#   print(tempVpdPlot)
# }

#Find appropriate linear models for each species/age
#PSMEo
# lmPSMEoVpd <- lm(PSMEo ~ Vpd, data = sapFluxEx)
# lmPSMEoRad <- lm(PSMEo ~ Rad, data = sapFluxEx)
lmPSMEoRadVpd <- lm(PSMEo ~ Rad + Vpd, data = sapFluxEx)

#PSMEy
# lmPSMEyVpd <- lm(PSMEy ~ Vpd, data = sapFluxEx)
# lmPSMEyRad <- lm(PSMEy ~ Rad, data = sapFluxEx)
lmPSMEyRadVpd <- lm(PSMEy ~ Rad + Vpd, data = sapFluxEx)

#TSHEo
# lmTSHEoVpd <- lm(TSHEo ~ Vpd, data = sapFluxEx)
# lmTSHEoRad <- lm(TSHEo ~ Rad, data = sapFluxEx)
lmTSHEoRadVpd <- lm(TSHEo ~ Rad + Vpd, data = sapFluxEx)

# summary(lmPSMEoVpd)
# summary(lmPSMEoRad)
# summary(lmPSMEoRadVpd)
# summary(lmPSMEyVpd)
# summary(lmPSMEyRad)
# summary(lmPSMEyRadVpd)
# summary(lmTSHEoVpd)
# summary(lmTSHEoRad)
# summary(lmTSHEoRadVpd)

#Model using both vpd and rad best option for all 3
sapFluxPred <- sapFluxEx %>% 
  add_predictions(model = lmPSMEoRadVpd, var = "PSMEoPred") %>% 
  add_predictions(model = lmPSMEyRadVpd, var = "PSMEyPred") %>% 
  add_predictions(model = lmTSHEoRadVpd, var = "TSHEoPred") %>% 
  mutate(Extrap = ifelse(is.na(PSMEo), 1, 0)) %>% #To mark what data is extrapolated vs measured
  mutate(PSMEo = ifelse(is.na(PSMEo), PSMEoPred, PSMEo),
         PSMEy = ifelse(is.na(PSMEy), PSMEyPred, PSMEy),
         TSHEo = ifelse(is.na(TSHEo), TSHEoPred, TSHEo))

#Plot extrapolated data
PSMEyPredPlot <- sapFluxPred %>% 
  filter(between(DATE, as.Date("1999-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = PSMEy, color = Extrap))
print(PSMEyPredPlot)

PSMEoPredPlot <- sapFluxPred %>% 
  filter(between(DATE, as.Date("1999-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = PSMEo, color = Extrap))
print(PSMEoPredPlot)

TSHEoPredPlot <- sapFluxPred %>% 
  filter(between(DATE, as.Date("1999-04-01"), as.Date("2000-10-31"))) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = TSHEo, color = Extrap))
print(TSHEoPredPlot)

#Combine sap flux data from all four species/ages into a single df
sapFluxAllPred <- sapFluxPred %>% 
  select(DATE, PSMEo, PSMEy, TSHEo) %>% 
  left_join(.,ALRUExPred, by = "DATE") %>% 
  rename(ALRUy = sFlux) %>% 
  select(-c(Rad, Vpd, sFluxPred, Extrap))

#Plot all sapFlux time series from May 1 through August 30
sFluxPlot <- sapFluxAllPred %>% 
  filter(between(DATE, as.Date("2000-05-01"), as.Date("2000-08-31"))) %>% 
  gather(tree, sFlux, PSMEo, PSMEy, TSHEo, ALRUy) %>% 
  ggplot(aes(x = DATE, y = sFlux, color = tree)) + geom_line()
print(sFluxPlot)

```

Data for manuscript figure X (transpiration deficits)
  - Calculate transpiration from sap flux density values (sapFluxAllPred) and SBA for appropriate classes 
    - See eqn 2 and related text in Moore et al
  - For hillslope hypothesis:
    - Difference between transpiration for the whole watershed - riparian area:
      - (PSMEo + TSHEo) - PSMEy
    - Multiplied by ratio of hillslope area to total area
  - For riparian hypothesis:
    - Difference between transpiration for just riparian area:
      - (PSMEo + TSHEo) - ALRUy 
    - Multiplied by ratio of riparian area to total area
      
```{r TranspirationDeficit}
#Inputs
StartDate = as.Date("2000-05-01")
EndDate = as.Date("2000-08-30")

ATot = 96
ARip = 2.18
AMoore = 10.46
AHill = ATot - ARip
AHillLow = AMoore - ARip

Tree <- c("PSME", "TSHE", "ALRU", "Ang", "Gym")
Old <- c(5.5, 9.9, 0.2, 1.3, 15.9)
Young <- c(13.9, 0.1, 1.5, 7.7, 13.9)

#Restructre SBA
#Modify SBA from Moore to account for "distributing" trees to different subareas of Moore "riparian" area in young stand
SBAall <- data.frame(Tree, Old, Young) %>% 
  gather(Stand, SBA, Old, Young) %>% 
  unite(col = "TreeStand", c("Tree", "Stand"), sep = "") %>% 
  mutate(SBA = ifelse(TreeStand == "PSMEYoung", (SBA * AMoore/AHillLow), SBA)) %>% #Modified SBA for PSMEy
  mutate(SBA = ifelse(TreeStand == "AngYoung", (SBA * AMoore/ARip), SBA)) #Modified SBA for ALRUy (use AngYoung)

#Calculate daily transpiration numbers
TransDef <- sapFluxAllPred %>% 
  mutate(transPSMEo = PSMEo * SBAall$SBA[SBAall$TreeStand == "PSMEOld"] / 10000,
         transTSHEo = TSHEo * SBAall$SBA[SBAall$TreeStand == "TSHEOld"] / 10000,
         transPSMEy = PSMEy * SBAall$SBA[SBAall$TreeStand == "PSMEYoung"] / 10000,
         transALRUy = ALRUy * SBAall$SBA[SBAall$TreeStand == "AngYoung"] / 10000,
         transOtherGym = (PSMEo + TSHEo)/2 * SBAall$SBA[SBAall$TreeStand == "GymOld"] / 10000) %>% 
  mutate(transHillOld =  (transPSMEo + transTSHEo + transOtherGym) * AHill/ATot,
         transHillYoung = transPSMEy * AHill/ATot,
         transRipOld = (transPSMEo + transTSHEo + transOtherGym) * ARip/ATot,
         transRipYoung = transALRUy * ARip/ATot) %>% 
  mutate(transDefHill = transHillOld - transHillYoung,
         transDefRip = transRipOld - transRipYoung) %>% 
  mutate(transOld = transHillOld + transRipOld,
         transYoung = transHillYoung + transRipYoung) %>% 
  mutate(transDef = transOld - transYoung) %>% 
  filter(between(DATE, StartDate, EndDate)) %>% 
  arrange(DATE)

save(TransDef, file = "TransDef.RData")
```

Make some plots of transpiration deficits
```{r TranspirationDeficitPlots}
#Transpiration from all tree species
allTreesTransPlot <- TransDef %>% 
  select(DATE, transPSMEo, transPSMEy, transTSHEo, transALRUy) %>% 
  rename(PSMEo = transPSMEo,
         PSMEy = transPSMEy,
         TSHEo = transTSHEo,
         ALRUy = transALRUy) %>% 
  gather(Tree, Trans, PSMEo, PSMEy, TSHEo, ALRUy) %>% 
  ggplot(aes(x = DATE, y = Trans, color = Tree)) + geom_line()
print(allTreesTransPlot)

#Overall Transpiration Deficit
OverallTransDefPlot <- TransDef %>% 
  select(DATE, transOld, transYoung) %>% 
  rename(Old = transOld, Young = transYoung) %>% 
  gather(Stand, Trans, Old, Young) %>% 
  ggplot(aes(x = DATE, y = Trans, color = Stand)) + geom_line()
print(OverallTransDefPlot)

#Hillslope Transpiration Deficit
HillslopeTransDefPlot <- TransDef %>% 
  select(DATE, transHillOld, transHillYoung, transDefHill) %>% 
  gather(Stand, Trans, transHillOld, transHillYoung, transDefHill) %>% 
  ggplot(aes(x = DATE, y = Trans, color = Stand)) + geom_line()
print(HillslopeTransDefPlot)

#Riparian Transpiration Deficit
RiparianTransDefPlot <- TransDef %>% 
  select(DATE, transRipOld, transRipYoung, transDefRip) %>% 
  gather(Stand, Trans, transRipOld, transRipYoung, transDefRip) %>% 
  ggplot(aes(x = DATE, y = Trans, color = Stand)) + geom_line()
print(RiparianTransDefPlot)

#Print out deficit numbers for various 
sum(TransDef$transDefHill)
sum(TransDef$transDefHill[between(TransDef$DATE, as.Date("2000-05-16"), as.Date("2000-06-9"))])
sum(TransDef$transDefHill[between(TransDef$DATE, as.Date("2000-05-01"), as.Date("2000-07-10"))])
sum(TransDef$transDefRip)
sum(TransDef$transDef)
```
```{r AGU23 Trans Figure}
#Set colors
youngCol = "#4dac26"
oldCol = "#a6611a"
ripDefCol = "#01665e"

myCols = c(c("Young" = youngCol, "Old" = oldCol, "RipDef" = ripDefCol))

myLineWidth = 2
myTxtSz = 20

#Add OpenET estimates to TransDef and rescale OpenET estimates to hillslope and riparian areas (ha)
ATot = 96
ARip = 2.18
AHill = ATot - ARip

TransOET <- TransDef %>% 
  select(DATE, transHillOld, transHillYoung, transDefHill,
         transRipOld, transRipYoung, transDefRip) %>%
  mutate(ET_OpenET = 
    case_when(month(DATE) == 5 ~ WS1_All_OET_month$Mean_Month_ET[5],
              month(DATE) == 6 ~ WS1_All_OET_month$Mean_Month_ET[6],
              month(DATE) == 7 ~ WS1_All_OET_month$Mean_Month_ET[7],
              month(DATE) == 8 ~ WS1_All_OET_month$Mean_Month_ET[8],
              TRUE ~ 0)) %>% 
  mutate(ET_OET_Rip = ET_OpenET * ARip/ATot,
         ET_OET_Hill = ET_OpenET * AHill/ATot)

#Riparian Transpiration Deficit without deficit
AGUFig <- TransOET %>% 
  select(DATE, transRipOld, transRipYoung, transDefRip) %>% 
  ggplot() +
  geom_area(aes(x = DATE, y = transRipYoung, fill = "Young"), 
            alpha = 0.3) + 
  geom_area(aes(x = DATE, y = transRipOld, fill = "Old"), 
            alpha = 0.3) +
  geom_area(aes(x = DATE, y = transDefRip, fill = "RipDef"),
            alpha = 0.7) +
  theme_classic(base_size = myTxtSz) +
  theme(panel.border = element_rect(fill=NA)) +
  scale_fill_manual(name = "", values = myCols) +
  labs(x = "", y = bquote("Transpiration (mm" ~d^-1*")")) +
  scale_x_date(expand = c(0,0))
print(AGUFig)

# #Panel C: Overall Transpiration Deficit
# myCols = c(Rip = ripDefCol, Hill = hillDefCol)
# 
# Fig7c <- TransOET %>% 
#   select(DATE, transDefRip, transDefHill) %>% 
#   mutate(Rip = cumsum(transDefRip),
#          Hill = cumsum(transDefHill)) %>% 
#   select(-c(transDefRip, transDefHill)) %>% 
#   gather(Position, sumTrans, Rip, Hill) %>% 
#   ggplot(aes(x = DATE, y = sumTrans, fill = Position)) +
#   geom_area(alpha = 0.7) +
#   scale_fill_manual(name = "", values = myCols) +
#   theme_bw(base_size = myTxtSz) +
#   scale_x_date(expand = c(0,0)) +
#   labs(x = "", y = "\u03A3 Transpiration (mm)")
# print(OverallTransDefPlot)  

```






